### 2016-02-02 paper/paper.tex
Add introduction ending and 2x2 condensation section

\RequirePackage{amsmath}
\documentclass{llncs}

\usepackage{cmap}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}

\usepackage{tikz}

\usepackage[font=small]{caption}
\usepackage{subcaption}

\usepackage{hyperref}

\usepackage{listings}
\usepackage{xcolor}

\usepackage{xspace}

\tikzset{every picture/.append style={scale=0.5}}

\tikzstyle{cell}=[rectangle,draw=gray,dashed,thick, minimum size=1.05cm]
\tikzstyle{graybox}=[rectangle,thick, minimum size=0.4cm, fill=lightgray]
\tikzstyle{whitebox}=[rectangle,draw=black,thick, minimum size=0.5cm]

\lstset{
  language=c++,
  basicstyle=\ttfamily \scriptsize,
  lineskip={-1.5pt},
  columns=fixed,
  basewidth=0.5em,
  keywordstyle=\bfseries\color{blue!60!black},
  commentstyle=\itshape\color{gray!80!black},
  identifierstyle=\color{violet!60!black},
  stringstyle=\color{orange},
  frame=single
  % numbers=left
}

\renewcommand{\C}{\texttt{C}\xspace}

\begin{document}

\title{Optimization for connected-component counting in binary image for the case of
8-connectivity}

\author{Fedor Alekseev\inst{1} \and Mikhail Alekseev\inst{2}\inst{3}
\and Artyom Makovetskii\inst{2}\inst{4}}

\institute{Moscow Institute of Physics and Technology State University, Dolgoprudny, Russia\\
\email{alekseev@phystech.edu}
\and
Chelyabinsk State University, Chelyabinsk, Russia
\and
\email{alexeev@csu.ru}
\and
\email{artemmac@mail.ru}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
The linear variation is a topological characteristic of the function of two variables.
The problem of linear variation computing can be reduced to the problem
of counting connected components in a binary image with eight-connected connectivity.
The new hybrid method for the problem is presented.
The method is essentially a modification of a two-scan algorithm for labeling
that groups the pixels into $2 \times 2$ cells.

\keywords{Binary raster image, connected component labeling,
pattern recognition, parallel processing}
\end{abstract}

\section{Introduction}

This problem is equivalent to the problem of counting the number of connected
components (CC) in a binary image with 8-connectivity.
This is usually done through CC labeling, and many approaches for that are
known\cite{sterzh}.

This paper presents an efficient way to convert the initial binary image
with 8-connectivity into
a condensed (non-binary) image with new connectivity that is $2$ times smaller in
each dimension than the initial image.
Most approaches for CC labeling are still valid for the new matrix.
As time and memory consumptions of conventional approaches usually depend linearly on
the number of pixels, running them on the condensed image can be up to 4 times
more efficient than running on the initial image.

It is worth noting that our method exploits an important property of 8-connectivity
that does not hold for 4-connectivity, so it is not applicable for the latter case.

\section{$2\times2$ condensation}

We will use the fact that for the case of 8-connectivity all $4$ pixels of any
$2 \times 2$ square are adjacent to each other, and so are contained in the same CC
and should have same labels upon completion of the labeling algorithm.
So instead of labeling single pixels we can label $2 \times 2$ groups of pixels,
or ``cells''.

Consider a $2N \times 2M$ binary image\footnote{
For convenience in this paper we consider only input images with even size
in both dimensions. If this is not the case, one can easily append a row or
a column of background pixels to the image.}
specified by a binary predicate $b(x, y)$,
returning $1$ in case the pixel at the given coordinates is black (object pixel),
and $0$ otherwise (background pixel).
For convenience we use 0-indexation throughout the paper, so the topmost row has
row index $y=0$, and the leftmost column has column index $x=0$.

The condensed image will be of size $N \times M$.
Each pixel of the condensed image will unambiguously represent the configuration
of the corresponding $2\times 2$ cell of pixels of the initial binary image.
So the condensed image will contain pixels of $2^4 = 16$ colors.
We suggest the function $c(x, y)$ denoting the color of a pixel of condensed image
to be the following:
\begin{align*}
  c(x, y) &= 2^0 b(2x, 2y)   + 2^1 b(2x+1, 2y) \\
          &+ {} 2^2 b(2x, 2y+1) + 2^3 b(2x+1,2y+1)
\end{align*}

So the value of each pixel of the initial image is stored in the corresponding
bit in color of the corresponding pixel of the condensed image.
Note that if $c(x, y) = 0$ for some $(x, y)$, then there are no object pixels in
this cell, and no label should be assigned to this whole cell. The enumeration of
pixels inside one cell is shown on \autoref{fig:pixelsEnumeration}.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}
    \node[cell] at (0.5, 0.5) {};
    \foreach \x/\y [count=\i from 0] in {0/1,1/1,0/0,1/0} {
      \node[whitebox] at (\x,\y) {\i};
    }
  \end{tikzpicture}
  \caption{Cell pixels enumeration}
  \label{fig:pixelsEnumeration}
\end{figure}

We need also to define connectivity function for the condensed image.
We cannot actually just use 8-connectivity, as whether two cells are adjacent
cells of one CC depends not only on their coordinates, but also on their
configuration. See \autoref{fig:connectivity:examples} for examples.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,2.5}
        \node[cell] at (\x,0.5){};
      \foreach \x/\y in {0/0, 1/1, 3/0, 3/1}
        \node[whitebox] at (\x, \y){};
    \end{tikzpicture}
    \caption{Nearby cells that are not directly connected.}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,2.5}
        \node[cell] at (\x,0.5){};
      \foreach \x/\y in {0/0, 1/1, 2/0, 3/1}
        \node[whitebox] at (\x, \y){};
    \end{tikzpicture}
    \caption{Directly connected cells.}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.3\linewidth}
    \centering
    \begin{tikzpicture}
      \foreach \x in {0.5,4.5}
        \node[cell] at (\x,0.5){};
      \foreach \x/\y in {0/0, 1/1, 4/0, 5/1}
        \node[whitebox] at (\x, \y){};
    \end{tikzpicture}
    \caption{Cells that are not nearby in terms of 8-connectivity, cannot be
    directly connected.}
  \end{subfigure}
  \caption{Cells connectivity examples}
  \label{fig:connectivity:examples}.
\end{figure}

However, for every variant of relative placement of two nearby cells there is
a mask of pixels of each cells that are important for direct connectivity of
this pair of cells.
So two cells are considered directly connected, if they are nearby
in terms of 8-connectivity and each of them contains at least one object pixel
in the mask of important pixels corresponding to relative placement of the cells.
The masks are shown in \autoref{fig:connectivity:masks}.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 2.5) {};
      \node[cell] at (0.5, 0.5) {};
      \node[graybox] at (0, 2) {2};
      \node[graybox] at (1, 2) {3};
      \node[graybox] at (0, 1) {0};
      \node[graybox] at (1, 1) {1};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.2\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 2.5) {};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 2) {2};
    \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}[t]{0.25\linewidth}
    \centering
    \begin{tikzpicture}
      \node[cell] at (0.5, 0.5) {};
      \node[cell] at (2.5, 0.5) {};
      \node[graybox] at (1, 0) {3};
      \node[graybox] at (1, 1) {1};
      \node[graybox] at (2, 0) {2};
      \node[graybox] at (2, 1) {0};
    \end{tikzpicture}
  \end{subfigure}
  \caption{Masks of important pixels for all four cases of relative placement
  of two nearby cells. Those pixels are filled in gray, with their numbers within
  cell specified.}
  \label{fig:connectivity:masks}
\end{figure}

The idea is that if we naturally encode mask pixels as 4-bit numbers in the
way similar to how we defined $c$, we can express the predicate denoting whether
two nearby cells
are directly connected efficiently with just two bitwise and one logical AND
operations.
As an example, \autoref{fig:connectivity:code} compares possible codes in \C checking
if the labels of the current pixel or cell and the one right on top of it
should be same for classic 8-connectivity and for the new connectivity.
Note that as bitwise operations are usually relatively cheap, we added only a
little complication compared to decreasing the number of pixels to process by the 
factor of 4.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.475\linewidth}
    \centering
    \begin{lstlisting}
    if (b(x, y)) {
      if (b(x, y-1)) {
        // copy labels, or union label sets
      }
      // consider other directions
    }
    \end{lstlisting}
    \caption{Binary image with 8-connectivity.}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.475\linewidth}
    \centering
    \begin{lstlisting}
    if (c(x, y)) {
      if ((c(x, y-1) & ((1<<2) | (1<<3)))
       && (c(x, y)   & ((1<<0) | (1<<1)))) {
        // copy labels, or union label sets
      }
      // consider other directions
    }
    \end{lstlisting}
    \caption{Condensed image with new connectivity.}
  \end{subfigure}
  \caption{Checking the nearby top pixel/cell.}
  \label{fig:connectivity:code}
\end{figure}

\section{Benchmarks}

\begin{thebibliography}{5}

\bibitem {hechao}
He, L., Chao, Y., Suzuki, K., Wu, K.:
Fast connected-component labeling. Pattern Recognition. v42, Pages 1977-1987. (2009)

\bibitem {sterzh}
Стержанов, М.:
Методики выделения связных компонент в штриховых бинарных изображениях.
Труды конференции ГрафиКон’2010. НИУ ИТМО. Санкт-Петербург. 2010, с. 169-174.

\end{thebibliography}

\end{document}

### 2016-02-02 paper/paper.tex
Move a box in connectivity example
Start converting Makovetskii part to latex

@@
 
+\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
 \renewcommand{\C}{\texttt{C}\xspace}
@@
 \keywords{Binary raster image, connected component labeling,
-pattern recognition, parallel processing}
+pattern recognition}
 \end{abstract}
@@
 \section{Introduction}
+
+% Makovetskii part
+
+In image restoration it is often necessary to consider the following problem.
+How to restore an original undistorted image $v$, if it is known an observed
+image $u_0$ and the relation between $u_0$ and $v$:
+\begin{equation}
+  u_0=v+n,
+  \label{eq:imageRelation}
+\end{equation}
+where $n$ is a noise?
+A common way for solving the problems \ref{eq:imageRelation} is to use the
+variational functionals.
+One of the most widely used approaches is total variation 
+regularization [1]. % TODO fix bibliography
+Let $J(u)$ be the following functional:
+\begin{equation}
+  J(u) = \frac12 \norm{u - u_0}^2 + \lambda TV(u),
+  \label{eq:jFunctional}
+\end{equation}
+where $\norm{\cdot}$ is the $L_2$ norm and all function we consider belong to the class
+$BV(\Omega)$, i.\,e. the class of bounded variation on the set $\Omega$ functions.
+%
+The expression $\frac12 \norm{n-n_0}^2$ is called a fidelity term and
+$\lambda TV(u)$ is called a regularization term and $\lambda$ is regularization
+parameter.
+
+% Alekseev part
 
@@
         \node[cell] at (\x,0.5){};
-      \foreach \x/\y in {0/0, 1/1, 4/0, 5/1}
+      \foreach \x/\y in {0/0, 0/1, 1/1, 4/0, 4/1, 5/0}
         \node[whitebox] at (\x, \y){};

### 2016-02-02 paper/paper.tex
Almost finish Makovetskii part
All but bibliography and corrections.

@@
 \documentclass{llncs}
+
+\usepackage{amssymb}
 
@@
 \newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-\renewcommand{\C}{\texttt{C}\xspace}
+\newcommand*{\N}{\mathbb{N} \xspace}
+\DeclareMathOperator*{\argmin}{arg\,min}
+
+\renewcommand{\C}{\texttt{C} \xspace}
 
@@
 where $n$ is a noise?
-A common way for solving the problems \ref{eq:imageRelation} is to use the
+A common way for solving the problems \eqref{eq:imageRelation} is to use the
 variational functionals.
@@
 parameter.
+
+\begin{equation}
+  TV(u) = \int_{\Omega} |\nabla u|\,dx\,dy.
+  \label{eq:TV}
+\end{equation}
+
+Let $u_*$ be extremal function for the following variational problem:
+\begin{equation}
+  u_* = \argmin_{u \in BV(\Omega)} J(u).
+  \label{eq:ustar}
+\end{equation}
+The \autoref{eq:ustar} is called total variation regularization problem.
+
+The norm and modulus of the gradient are metrical characteristic of a function of
+two variables.
+Continuous functions of two variables also have a set of topological
+characteristics referred to as linear variations.
+Kronrod [2] % TODO fix bibliography
+introduced the notion of a regular component of the level set of a
+continuous function of two variables.
+The simplest topological characteristic in the linear variation theory is a number
+of regular components for all level sets of a function.
+Full information about linear variations of a function is contained in the
+one-dimensional tree of a function of two variables.
+In [3, 4, 5] % TODO fix bibliography
+was discussed the necessity of the using a linear variation in the
+image restoration theory.
+
+Let $\Phi_u(t)$ be the number of regular components of a level set $t$
+for a continuous function.
+The first Kronrod's linear variation is defined as
+\begin{equation}
+  V(u) = \int_{-\infty}^{+\infty} \Phi_u(t)\,dt.
+  \label{eq:KronrodVariation}
+\end{equation}
+
+Let $w$ be a binary discrete function $w=(w_{i,j})$, where $w_{i,j} \in \{0,1\}$,
+for all pairs $(i,j)$.
+A subset of such pairs $(i, j)$ when $w_{i,j}=1$ and all elements of the subset are
+connected by the 8-connectivity, is called the connected component of the binary
+function $w$.
+For a number $k \in \N$ and a discrete function $u$ we define the following
+indicator function $\chi$:
+\begin{equation}
+  \chi_k (u_{i,j}) =
+    \begin{cases}
+      1, & u_{i,j} \ge k \\
+      0, & u_{i,j} < k
+    \end{cases}.
+  \label{eq:chiIndicator}
+\end{equation}
+
+\begin{definition}
+  The number $V_k(u_{i,j})$ of connected components for a level $k$, $k \in \N$ of
+  the discrete function $u$ is called the number of connected components of the 
+  binary discrete function $\chi_k(u_{i,j})$.
+\end{definition}
+
+\begin{definition}
+  The linear variation $V(u_{i,j})$ of a discrete function  is defined as follows:
+  \begin{equation}
+    V(u_{i,j}) = \sum_{k=0}^{+\infty} V_k(u_{i,j})
+    \label{eq:V}
+  \end{equation}
+\end{definition}
+
+Let us compute the discrete gradient $\nabla u_{i,j}$ of $u$ at $(i,j)$ as
+\begin{equation}
+  \nabla u_{i,j} = (u_{i+1,j} - u_{i,j}, u_{i,j+1} - u_{i,j}).
+  \label{eq:gradient}
+\end{equation}
+
+Suppose that if the pair $(i,j)$ is outside of the domain of the function $u$, then 
+$u_{i,j} = 0$.
+The main problem of the computer implementation of the linear variation is 
+computation of $V_k(u_{i,j})$ for given k.
 

### 2016-02-02 paper/paper.tex
Fix title and integral limits placement

@@
 
-\title{Optimization for connected-component counting in binary image for the case of
-8-connectivity}
+\title{Linear variation and an optimization of algorithms for connected
+components labeling in a binary image}
 
@@
 \begin{abstract}
-The linear variation is a topological characteristic of the function of two variables.
+Linear variation is a topological characteristic of a function of two variables.
 The problem of linear variation computing can be reduced to the problem
@@
 \begin{equation}
-  TV(u) = \int_{\Omega} |\nabla u|\,dx\,dy.
+  TV(u) = \int\limits_{\Omega} |\nabla u|\,dx\,dy.
   \label{eq:TV}
@@
 \begin{equation}
-  V(u) = \int_{-\infty}^{+\infty} \Phi_u(t)\,dt.
+  V(u) = \int\limits_{-\infty}^{+\infty} \Phi_u(t)\,dt.
   \label{eq:KronrodVariation}
@@
 each dimension than the initial image.
-Most approaches for CC labeling are still valid for the new matrix.
+Most approaches for CC labeling are still valid for the new image.
 As time and memory consumptions of conventional approaches usually depend linearly on

### 2016-02-05 utils/csv-to-latex.py
Add benchmark tables for several sizes

@@ -1,0 +1,65 @@
+#!/usr/bin/env python3
+
+import sys
+import json
+import csv
+import re
+import itertools
+import os
+import collections
+
+
+def read_benchmarks_csv(csv_file=None):
+    csv_file = csv_file or sys.stdin
+    reader = csv.DictReader(csv_file)
+    return list(reader)
+
+
+COUNTER_MAP = {
+    "Dfs": "DFS",
+    "Dsu": "SUF",
+}
+
+
+def write_benchmarks_latex(runs):
+    runs.sort(key=lambda r: (r['counter'], '2x2' in r['grid'], int(r['density'])))
+
+    headers = ['counter', '2x2']
+    densities = set()
+    lines = []
+
+    for (counter, cond), runs in itertools.groupby(runs, lambda r: (r['counter'], '2x2' in r['grid'])):
+        runs = tuple(runs)
+        counter = counter[:counter.index('Counter')]
+        line = {
+            headers[0]: COUNTER_MAP.get(counter, counter),
+            headers[1]: 'yes' if cond else 'no',
+        }
+        for r in runs:
+            dens = r['density']
+            densities.add(dens)
+            line[dens] = str(round(float(r['cpu time ms']), 1))
+        lines.append(line)
+
+    columns = len(headers) + len(densities)
+    print(r'\begin{tabular}{| l | r | ' + 'r ' * (columns - 2) + '| }')
+
+    print(r'\cline{' + '{}-{}'.format(3, columns) + '}')
+    print(r'\multicolumn{' + str(len(headers)) + '}{c}{} & '
+        + r'\multicolumn{' + str(len(densities)) + '}{|c|}{density}', end='\\\\\n')
+
+    print('\\hline')
+    headers.extend(map(str, sorted(map(int, densities))))
+    print(' & '.join(headers), end='\\\\\n')
+    print('\\hline')
+    for line in lines:
+        print(' & '.join([line.get(k, '') for k in headers]), end='\\\\\n')
+
+    print('\\hline')
+    print(r'\end{tabular}')
+
+
+if __name__ == '__main__':
+    runs = read_benchmarks_csv()
+    write_benchmarks_latex(runs)
+

### 2016-02-05 utils/report.py

@@ -4,6 +4,8 @@ import sys
 import json
 import csv
 import re
+import itertools
+import os
 
 NAME_RE = re.compile(
         '^BM<' +
@@ -11,7 +13,8 @@ NAME_RE = re.compile(
         ',\s*' +
         '(?P<test_signature>'
         '(?P<type>\w+?)_(?P<matrix>\w+?)(_(?P<height>\d+)x(?P<width>\d+))?' +
-        '(_(?P<id>(\d|\w|_)*))?' +
+        '(_(?P<id>\d+))?' +
+        '(_(?P<density>\d+))?'
         ')>$')
 
 TIME_MULTIPLIER = 1e-6
@@ -31,7 +34,7 @@ def write_benchmarks_csv(runs, csv_file=None):
     csv_file = csv_file or sys.stdout
     writer = csv.DictWriter(csv_file,
             ['type', 'counter', 'params', 'grid', 'matrix',
-             'height', 'width', 'id', 'cpu time ms'],
+             'height', 'width', 'id', 'density', 'cpu time ms'],
             extrasaction='ignore')
     writer.writeheader()
     for r in runs:
@@ -64,9 +67,26 @@ def update_best(new_bests, filepath):
     with open(filepath, 'w') as fout:
         json.dump(old_bests, fout, sort_keys=True, indent=2)
 
+def path_for_size(h, w):
+    return os.path.join('bench-out', '{}x{}.csv'.format(h, w))
+
+def split_by_sizes(runs):
+    def key(r):
+        return (r['height'], r['width'])
+    runs.sort(key=key)
+    return [((h, w), list(r))
+            for (h, w), r
+            in itertools.groupby(runs, key=key)]
+
 if __name__ == '__main__':
     runs = read_benchmarks_json()
     if '--update-best' in sys.argv:
         update_best(choose_best(runs), 'best.json')
-    write_benchmarks_csv(runs)
+    if '--split-by-sizes' in sys.argv:
+        for (h, w), sub_runs in split_by_sizes(runs):
+            #  print(sub_runs[0], file=sys.stderr)
+            with open(path_for_size(h, w), 'w') as outfile:
+                write_benchmarks_csv(sub_runs, outfile)
+    else:
+        write_benchmarks_csv(runs)
 

### 2016-02-05 utils/csv-to-latex.py
Add viewing grid, refactor grids a bit

@@ -22,18 +22,20 @@ COUNTER_MAP = {
 
 
 def write_benchmarks_latex(runs):
-    runs.sort(key=lambda r: (r['counter'], '2x2' in r['grid'], int(r['density'])))
+    grids = ['no', 'view', 'in-memory']
+    gridfunc = lambda g: 1 if 'View' in g else 2 if '2x2' in g else 0
+    runs.sort(key=lambda r: (r['counter'], gridfunc(r['grid']), int(r['density'])))
 
-    headers = ['counter', '2x2']
+    headers = ['method', 'compression']
     densities = set()
     lines = []
 
-    for (counter, cond), runs in itertools.groupby(runs, lambda r: (r['counter'], '2x2' in r['grid'])):
+    for (counter, compression), runs in itertools.groupby(runs, lambda r: (r['counter'], gridfunc(r['grid']))):
         runs = tuple(runs)
         counter = counter[:counter.index('Counter')]
         line = {
             headers[0]: COUNTER_MAP.get(counter, counter),
-            headers[1]: 'yes' if cond else 'no',
+            headers[1]: grids[compression],
         }
         for r in runs:
             dens = r['density']

### 2016-02-06 utils/csv-to-latex.py
Return TwolineDsuCounter back, some polish

@@ -16,23 +16,24 @@ def read_benchmarks_csv(csv_file=None):
 
 
 COUNTER_MAP = {
-    "Dfs": "DFS",
-    "Dsu": "SUF",
+    "DesisionTreeDsuCounter": "WuOtoo",
+    "DfsCounter": "DFS",
+    "DsuCounter": "SUF",
+    "TwolineDsuCounter": "SUF2"
 }
 
 
 def write_benchmarks_latex(runs):
-    grids = ['no', 'view', 'in-memory']
+    grids = ['no', 'view', 'yes']
     gridfunc = lambda g: 1 if 'View' in g else 2 if '2x2' in g else 0
     runs.sort(key=lambda r: (r['counter'], gridfunc(r['grid']), int(r['density'])))
 
-    headers = ['method', 'compression']
+    headers = ['method', '2x2']
     densities = set()
     lines = []
 
     for (counter, compression), runs in itertools.groupby(runs, lambda r: (r['counter'], gridfunc(r['grid']))):
         runs = tuple(runs)
-        counter = counter[:counter.index('Counter')]
         line = {
             headers[0]: COUNTER_MAP.get(counter, counter),
             headers[1]: grids[compression],

### 2016-02-06 utils/csv-to-latex.py
Some more polish for tables

@@ -16,7 +16,7 @@ def read_benchmarks_csv(csv_file=None):
 
 
 COUNTER_MAP = {
-    "DesisionTreeDsuCounter": "WuOtoo",
+    "DesisionTreeDsuCounter": "DTSUF",
     "DfsCounter": "DFS",
     "DsuCounter": "SUF",
     "TwolineDsuCounter": "SUF2"
@@ -26,7 +26,11 @@ COUNTER_MAP = {
 def write_benchmarks_latex(runs):
     grids = ['no', 'view', 'yes']
     gridfunc = lambda g: 1 if 'View' in g else 2 if '2x2' in g else 0
-    runs.sort(key=lambda r: (r['counter'], gridfunc(r['grid']), int(r['density'])))
+    runs.sort(key=lambda r: (
+        'Desis' in r['counter'],
+        r['counter'],
+        gridfunc(r['grid']),
+        int(r['density'])))
 
     headers = ['method', '2x2']
     densities = set()
@@ -54,11 +58,12 @@ def write_benchmarks_latex(runs):
     print('\\hline')
     headers.extend(map(str, sorted(map(int, densities))))
     print(' & '.join(headers), end='\\\\\n')
-    print('\\hline')
-    for line in lines:
-        print(' & '.join([line.get(k, '') for k in headers]), end='\\\\\n')
+    print('\\hline \\hline')
+    for _, counter_lines in itertools.groupby(lines, lambda li: li[headers[0]]):
+        for line in counter_lines:
+            print(' & '.join([line.get(k, '') for k in headers]), end='\\\\\n')
+        print('\\hline')
 
-    print('\\hline')
     print(r'\end{tabular}')
 
 

### 2016-02-06 paper/paper.tex
Add benchmark tables for several sizes
Return TwolineDsuCounter back, some polish
Some more polish for tables
Add overview for DFS and start one for SUF

@@
 
+In this section we present performance benchmarks of some popular methods for
+counting CC both with and without the $2 \times 2$ cells optimization.
+
+\subsection{Algorithms overview}
+
+\subsubsection{DFS}
+
+The first algorithm we chose for the benchmark is Depth First Search graph traversal.
+Graph traversal algorithms are the general approach for CC labeling in graphs, and
+DFS is arguably the most simple.
+Although not intended for image labeling and
+suboptimal in most cases, it still has linear in the number of pixels worst-case
+complexity in both time and memory.
+
+Like most approaches, the algorithm contains has and outer loop over all pixels
+of the image.
+Each time an object pixel that is not labeled yet is encountered, DFS traversal
+is launched from that pixel.
+The traversal labels all the object pixels of the same CC, launching recursively
+from all unlabeled neighbors of current pixel.
+
+\subsubsection{SUF}
+
+Another general approach for graph CC labeling which is more popular in image
+labeling is using some kind of a data structure to maintain disjoint sets of pixels
+that are considered to be in the same CC.
+The structure is commonly called \texttt{Union-Find}, as the two essential operations
+it must support are \texttt{Union}, which is used to union some of the disjoint sets,
+and \texttt{Find}, which is used to find the representative element of the set
+some element belongs to, so that we can always tell if two elements are in the same
+set.
+
+\subsubsection{SUF2}
+
+\subsubsection{DTSUF}
+
+\newcommand{\inputtable}[2]{
+  \begin{table}[p]
+    \centering
+    \input{../bench-out/#1x#2}
+    \caption{Noise images $#1 \times #2$ px}
+    \label{tab:noise#1x#2}
+  \end{table} }
+
+% \inputtable{200}{300}
+% \inputtable{400}{600}
+\inputtable{800}{1200}
+% \inputtable{1200}{1800}
+\inputtable{2000}{3000}
+\inputtable{4000}{6000}
+
 \begin{thebibliography}{5}

### 2016-02-06 paper/paper.tex
Finish SUF subsubsubsubsubsection

@@
 
+As each of these methods, omitting DTSUF could be used with both condensed and
+non-condensed images, for convenience we denote height and width of the image
+on input of each algorithm as $H$ and $W$.
+
+Also to avoid repeating ``pixel or cell'' we will just use the word ``pixel''
+meaning the appropriate raster for non-condensed and condensed images.
+When we mention neighbors of some pixel, we mean pixels that are directly connected
+to it in terms of appropriate connectivity.
+
 \subsubsection{DFS}
@@
 
-Like most approaches, the algorithm contains has and outer loop over all pixels
-of the image.
-Each time an object pixel that is not labeled yet is encountered, DFS traversal
-is launched from that pixel.
+Like most approaches, the algorithm has an outer loop over all pixels of the image.
+Each time an object pixel (non-empty cell, in case it is actually a cell)
+which is not labeled yet is encountered, DFS traversal is launched from that pixel.
 The traversal labels all the object pixels of the same CC, launching recursively
@@
 that are considered to be in the same CC.
-The structure is commonly called \texttt{Union-Find}, as the two essential operations
-it must support are \texttt{Union}, which is used to union some of the disjoint sets,
-and \texttt{Find}, which is used to find the representative element of the set
-some element belongs to, so that we can always tell if two elements are in the same
-set.
+
+The structure is commonly called Union-Find, as the two essential operations
+it is supposed to support are Union, which is used to merge some two of the
+disjoint sets in one, and Find, which is used to find the representative element
+of the set some element belongs to, so that we can always tell if two elements are
+in the same set.
+The structure can be implemented so that the amortized time complexity for each
+operation is $O(\alpha(HW))$, where $\alpha$ is the inverse Ackermann function,
+and $\alpha(HW) < 5$ for all remotely practical image sizes. % TODO links
+
+The approach that is named SUF (Scan plus Union-Find) in this paper follows.
+
+Pixels are processed in the natural order: from top to bottom and from
+left to right, so each row is fully processed before all rows after it.
+So for each object pixel we encounter there are at most 4 nearby pixels
+that were already processed, see \autoref{fig:suf:neighbors}.
+
+At first each object pixel is considered to belong to its own CC, so right before
+considering the neighbors we increment the CC counter.
+Moreover, it is associated with its own disjoint set in the Union-Find structure.
+More specifically, object pixel at position $x, y$ is associated with
+the set number $xW + y$ to guarantee absence of collisions.
+
+The possible neighbors are processed in the order shown on \autoref{fig:suf:neighbors},
+and for each neighbor that is object pixel, the Union operation is performed
+on the two sets containing the current and the neighbor pixel.
+In case the Union operation was successful, which means that the pixels
+had belonged previously to different sets and we have just found a way to merge two
+CC, the CC counter is decremented.
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \foreach \x/\y [count=\i from 1] in {0/1,1/1,2/1,0/0} {
+      \node[whitebox] at (\x,\y) {\i};
+    }
+    \node[graybox] at (1, 0) {};
+    \node[whitebox] at (1, 0) {};
+  \end{tikzpicture}
+  \caption{Neighbor mask for SUF algorithm family.}
+  \label{fig:suf:neighbors}
+\end{figure}
+
+Note that this algorithm outputs not the matrix of final labels, but only the number
+of CC.
+This is the place in which it differs from some common two-pass approaches for CC % TODO references
+labeling, as it is an algorithm not for labeling, but just for counting CC.
+The latter is exactly what is needed for the aforementioned linear variation problem.
+
+Also we have to mention that for performance reasons the Union-Find internal structures
+have to be allocated ahead to have the size of at least $HW$ to guarantee that
+each object can be easily associated with unique starting set given only its
+coordinates. So memory consumption of SUF is linear in the number of pixels.
 
@@
 \subsubsection{DTSUF}
+
+using desision trees suggested by Wu, Otoo and Suzuki
+
+\subsection{Condensation types}
+
+bla-bla about 2x2 in memory and as a view
 

### 2016-02-06 utils/csv-to-latex.py
Desision -> Decision

@@ -16,7 +16,7 @@ def read_benchmarks_csv(csv_file=None):
 
 
 COUNTER_MAP = {
-    "DesisionTreeDsuCounter": "DTSUF",
+    "DecisionTreeDsuCounter": "DTSUF",
     "DfsCounter": "DFS",
     "DsuCounter": "SUF",
     "TwolineDsuCounter": "SUF2"

### 2016-02-06 paper/paper.tex
Finish algorithms description, including the decision tree

@@
 \usepackage{hyperref}
+\usepackage{cleveref}
 
@@
 
-\tikzstyle{cell}=[rectangle,draw=gray,dashed,thick, minimum size=1.05cm]
+\tikzstyle{cell}=[rectangle,draw=gray,dashed,semithick, minimum size=1.05cm]
 \tikzstyle{graybox}=[rectangle,thick, minimum size=0.4cm, fill=lightgray]
 \tikzstyle{whitebox}=[rectangle,draw=black,thick, minimum size=0.5cm]
+
+% \captionsetup[subfigure]{subrefformat=simple,labelformat=simple}
+% \renewcommand\thesubfigure{(\alph{subfigure})}
 
@@
 As time and memory consumptions of conventional approaches usually depend linearly on
-the number of pixels, running them on the condensed image can be up to 4 times
+the number of pixels, running them on the condensed image can be
 more efficient than running on the initial image.
@@
 Also we have to mention that for performance reasons the Union-Find internal structures
-have to be allocated ahead to have the size of at least $HW$ to guarantee that
+are allocated ahead to have the size of at least $HW$ to guarantee that
 each object can be easily associated with unique starting set given only its
@@
 
+This approach is a modification of SUF designed to reduce memory consumption
+and potential number of cache-misses generated by Union-Find operations.
+
+The idea is to run through each row twice: first time to establish label equality
+classes within row, and the second time to set unique equal label for all pixels of
+same CC.
+This gives us the opportunity to maintain just $O(W)$
+different labels at each moment, and so the Union-Find structure can be allocated
+for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.
+
+The exact maximum number of label equality classes between two adjacent rows
+depends on whether the given image is condensed.
+For non-condensed images $W=2M$, and maximum number of label equality classes
+in two adjacent rows is $M=\frac{W}2$, as two object pixels are both in adjacent
+rows and in adjacent columns, then they are neighbors in terms of 8-connectivity.
+
+On the other hand, for condensed images $W=M$, and maximum number of label equality
+class in two adjacent rows of cells is $2W=2M$.
+See \autoref{fig:suf2:extremal} for illustration on the extremal cases.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{0.475\linewidth}
+    \centering
+    \begin{tikzpicture}
+      \foreach \x in {0,1,...,9} \foreach \y in {0,1} {
+        \node[whitebox,dashed,semithick,draw=gray] at (\x,\y) {};
+      }
+      \foreach \x/\y in {0/0,2/1,4/1,6/0,8/1} {
+        \node[whitebox] at (\x,\y) {};
+      }
+    \end{tikzpicture}
+    \caption{Non-condensed image, up to 5 different labels on 10 columns.}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}{0.475\linewidth}
+    \centering
+    \begin{tikzpicture}
+      \foreach \x in {0.5,2.5,...,8.5} \foreach \y in {0.5,2.5} {
+        \node[cell] at (\x,\y) {};
+      }
+      \foreach \x in {0,2,...,8} \foreach \y in {0,3} {
+        \node[whitebox] at (\x,\y) {};
+      }
+    \end{tikzpicture}
+    \caption{Condensed image, up to 10 different labels on 5 columns.}
+  \end{subfigure}
+  \caption{Extremal cases for SUF2.}
+  \label{fig:suf2:extremal}
+\end{figure}
+
+Our benchmarks show that, although SUF2 without condensation is generally more
+efficient on noise images than SUF without condensation, the mentioned
+consideration make the $2 \times 2$ optimization
+of SUF2 less dramatic in means of performance increase factor.
+Optimized SUF2 significantly outperforms optimized SUF only on
+huge noise images with less than 60\% density, while falling behind on
+small and medium-sized noise images.
+
+Note that, while SUF2 is intended for CC counting, its memory
+consumption is only $O(W)$, which is significant for embedded systems and
+other resource-tight applications.
+
 \subsubsection{DTSUF}
 
-using desision trees suggested by Wu, Otoo and Suzuki
+This modification of SUF was taken to compare the optimization we present with
+another optimization suggested by Wu, Otoo and Suzuki, % TODO reference
+namely decision tree.
+Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
+applicable for condensed image, as it exploits similar observations on 8-connectivity.
+
+As before, the goal is to reduce the number of pixel lookup and label copy operations.
+This is achieved with careful consideration of only the necessary neighbors,
+stopping once the set of one or two label operations that are needed is
+determined. The decision tree for processing an object pixel is
+shown on \autoref{fig:dtsuf}.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{0.3\linewidth}
+    \centering
+    \begin{tikzpicture}
+      \node[graybox] at (1,0) {};
+      \foreach \x/\y/\a in {0/1/a,1/1/b,2/1/c,0/0/d,1/0/e} {
+        \node[whitebox,font=\bf] at (\x,\y) {\a};
+      }
+    \end{tikzpicture}
+    \caption{Neighbor enumeration.}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}{0.65\linewidth}
+    \centering
+    \begin{tikzpicture}[
+      sibling distance=18em,
+      lookup/.style = {shape=circle, draw, thick, align=center, font=\bf},
+      action/.style = {shape=rectangle, rounded corners, draw, align=center, font=\scriptsize},
+      result/.style = {font=\bf}
+    ]
+      \node[lookup] {b}
+        child[level distance=3em,sibling distance=30em] {
+          node[lookup] {c}
+          child[level distance=5em, sibling distance=14em] {
+            node[lookup] {a}
+            child[level distance=6em, sibling distance=10em] {
+              node[lookup] {d}
+              child {
+                node[action] {new label}
+                edge from parent node[result,above] {$0$}
+              }
+              child {
+                node[action] {union(e,d)}
+                edge from parent node[result,above] {$1$}
+              }
+              edge from parent node[result,above] {$0$}
+            }
+            child {
+              node[action] {union(e,a)}
+              edge from parent node[result,above] {$1$}
+            }
+            edge from parent node[result,above] {$0$}
+          }
+          child[level distance=9em,sibling distance=24em] {
+            node[lookup] {a}
+            child[level distance= 6em, sibling distance=10em] {
+              node[lookup] {d}
+              child {
+                node[action] {union(e,c)}
+                edge from parent node[result,above] {$0$}
+              }
+              child {
+                node[action] {union(e,c,d)}
+                edge from parent node[result,above] {$1$}
+              }
+              edge from parent node[result,above] {$0$}
+            }
+            child[sibling distance=20em] {
+              node[action] {union(e,c,a)}
+              edge from parent node[result,above] {$1$}
+            }
+            edge from parent node[result,above] {$1$}
+          }
+          edge from parent node[result,above] {$0$}
+        }
+        child[level distance=2em] {
+          node[action] {union(e,b)}
+          edge from parent node[result,above] {$1$}
+        };
+    \end{tikzpicture}
+    \caption{The decision tree.}
+  \end{subfigure}
+  \caption{DTSUF decision tree optimization.}
+  \label{fig:dtsuf}
+\end{figure}
 

### 2016-02-06 paper/paper.tex
Add condensation types subsubsubsection

@@
 
-This modification of SUF was taken to compare the optimization we present with
+This modification of SUF was implemented to compare the optimization we present with
 another optimization suggested by Wu, Otoo and Suzuki, % TODO reference
@@
 
-bla-bla about 2x2 in memory and as a view
+For memory-tight applications such as embedded systems it can be useful to
+utilise the suggested condensation in a lazy way, so that without actually
+constructing the condensed image we use a function (``view'') calculating color of
+each cell with 4 lookups to the original binary image.
+This is probably most interesting combined with SUF2 algorithm, as it is a way
+to win some time without increasing memory footprint.
+However this lazy approach increases the number of image lookups.
+So, according to our benchmark, this combination is efficient only for dense
+($>30\%$) noise.
+
+So the benchmark tables show three levels of optimization applied to
+DFS, SUF and SUF2:
+``no'' meaning no condensation, ``view'' meaning lazy memory-less condensation
+and ``yes'' meaning that a temporary condensed image was constructed in memory.
+For the latter case, the construction time was included into the overall measured
+time of run.
+
+\subsection{Results}
+
+some results
+
+\section{Summary and future work}
 
@@
     \input{../bench-out/#1x#2}
-    \caption{Noise images $#1 \times #2$ px}
+    \caption{Average milliseconds of cpu time per run on noise images $#1 \times #2$ px}
     \label{tab:noise#1x#2}

### 2016-02-06 utils/csv-to-latex.py

@@ -27,7 +27,7 @@ def write_benchmarks_latex(runs):
     grids = ['no', 'view', 'yes']
     gridfunc = lambda g: 1 if 'View' in g else 2 if '2x2' in g else 0
     runs.sort(key=lambda r: (
-        'Desis' in r['counter'],
+        'Decis' in r['counter'],
         r['counter'],
         gridfunc(r['grid']),
         int(r['density'])))

### 2016-02-06 utils/csv-to-latex.py
Max out tables width

@@ -9,6 +9,14 @@ import os
 import collections
 
 
+def my_round(x):
+    from math import log10, floor
+
+    x = float(x)
+    x = round(x, -int(floor(log10(abs(x)))) + 1)
+    return round(x) if x - round(x) < 1e-4 and x >= 10 else x
+
+
 def read_benchmarks_csv(csv_file=None):
     csv_file = csv_file or sys.stdin
     reader = csv.DictReader(csv_file)
@@ -21,7 +29,6 @@ COUNTER_MAP = {
     "DsuCounter": "SUF",
     "TwolineDsuCounter": "SUF2"
 }
-
 
 def write_benchmarks_latex(runs):
     grids = ['no', 'view', 'yes']
@@ -45,11 +52,12 @@ def write_benchmarks_latex(runs):
         for r in runs:
             dens = r['density']
             densities.add(dens)
-            line[dens] = str(round(float(r['cpu time ms']), 1))
+            line[dens] = str(my_round(r['cpu time ms']))
         lines.append(line)
 
     columns = len(headers) + len(densities)
-    print(r'\begin{tabular}{| l | r | ' + 'r ' * (columns - 2) + '| }')
+    print(r'\begin{tabularx}{\textwidth}{| l  r  @{\extracolsep{\fill}}'
+            + 'r ' * (columns - 2) + '| }')
 
     print(r'\cline{' + '{}-{}'.format(3, columns) + '}')
     print(r'\multicolumn{' + str(len(headers)) + '}{c}{} & '
@@ -58,13 +66,13 @@ def write_benchmarks_latex(runs):
     print('\\hline')
     headers.extend(map(str, sorted(map(int, densities))))
     print(' & '.join(headers), end='\\\\\n')
-    print('\\hline \\hline')
+    print('\\hline')
     for _, counter_lines in itertools.groupby(lines, lambda li: li[headers[0]]):
         for line in counter_lines:
             print(' & '.join([line.get(k, '') for k in headers]), end='\\\\\n')
         print('\\hline')
 
-    print(r'\end{tabular}')
+    print(r'\end{tabularx}')
 
 
 if __name__ == '__main__':

### 2016-02-06 utils/csv-to-latex.py
Try to make some cool charts

@@ -30,7 +30,7 @@ COUNTER_MAP = {
     "TwolineDsuCounter": "SUF2"
 }
 
-def write_benchmarks_latex(runs):
+def regroup_runs(runs):
     grids = ['no', 'view', 'yes']
     gridfunc = lambda g: 1 if 'View' in g else 2 if '2x2' in g else 0
     runs.sort(key=lambda r: (
@@ -39,21 +39,54 @@ def write_benchmarks_latex(runs):
         gridfunc(r['grid']),
         int(r['density'])))
 
-    headers = ['method', '2x2']
     densities = set()
     lines = []
 
     for (counter, compression), runs in itertools.groupby(runs, lambda r: (r['counter'], gridfunc(r['grid']))):
         runs = tuple(runs)
         line = {
-            headers[0]: COUNTER_MAP.get(counter, counter),
-            headers[1]: grids[compression],
+            'method': COUNTER_MAP.get(counter, counter),
+            '2x2': grids[compression],
         }
         for r in runs:
-            dens = r['density']
+            dens = int(r['density'])
             densities.add(dens)
-            line[dens] = str(my_round(r['cpu time ms']))
+            line[dens] = float(r['cpu time ms'])
         lines.append(line)
+    return lines, sorted(densities)
+
+
+METHOD_COLOR = {
+    'DFS': 's',
+    'SUF': 'o',
+    "SUF2": '^',
+    'DTSUF': '*'
+}
+
+COND_SHAPE = {
+    'no': 'r',
+    'view': 'y',
+    'yes': 'g'
+}
+
+def plot_benchmarks(runs, out_file='output.eps'):
+    from matplotlib import pyplot as plt
+
+    lines, densities = regroup_runs(runs)
+
+    for line in lines:
+        method = line['method']
+        compression = line['2x2']
+        ys = [line[x] for x in densities]
+        plt.plot(densities, ys, COND_SHAPE[compression] + METHOD_COLOR[method] + '-.')
+    plt.xlabel('Noise density')
+    plt.ylabel('ms')
+    plt.savefig(out_file, format='eps', dpi=1200)
+
+def write_benchmarks_latex(runs):
+    headers = ['method', '2x2']
+    grids = ['no', 'view', 'yes']
+    lines, densities = regroup_runs(runs)
 
     columns = len(headers) + len(densities)
     print(r'\begin{tabularx}{\textwidth}{| l  r  @{\extracolsep{\fill}}'
[…]

### 2016-02-06 paper/paper.tex
Add 5000x7500 test
Max out tables width
Fix title capitalization
Add pictures instead of tables!

@@
 \usepackage{xspace}
+
+\usepackage{tabularx}
+\usepackage{graphicx}
 
@@
 
-\title{Linear variation and an optimization of algorithms for connected
-components labeling in a binary image}
+\title{Linear Variation and an Optimization of Algorithms for Connected
+Components Labeling in a Binary Image}
 
@@
 
-\section{Summary and future work}
-
 \newcommand{\inputtable}[2]{
-  \begin{table}[p]
+  \begin{figure}
     \centering
-    \input{../bench-out/#1x#2}
-    \caption{Average milliseconds of cpu time per run on noise images $#1 \times #2$ px}
-    \label{tab:noise#1x#2}
-  \end{table} }
+    \includegraphics[height=0.45\textheight, keepaspectratio]{pics/#1x#2.eps}
+    \caption{Average CPU time per run on noise images $#1 \times #2$ px}
+    \label{fig:noise#1x#2}
+  \end{figure}
+  % \begin{table}[p]
+    % \centering
+    % \input{../bench-out/#1x#2}
+    % \caption{Average milliseconds of cpu time per run on noise images $#1 \times #2$ px}
+    % \label{tab:noise#1x#2}
+  % \end{table}
+}
 
@@
 \inputtable{4000}{6000}
+\inputtable{5000}{7500}
+
+\section{Summary and future work}
 

### 2016-02-06 utils/csv-to-latex.py

@@ -55,33 +55,54 @@ def regroup_runs(runs):
         lines.append(line)
     return lines, sorted(densities)
 
-
 METHOD_COLOR = {
-    'DFS': 's',
-    'SUF': 'o',
-    "SUF2": '^',
-    'DTSUF': '*'
+    'DFS': 'red',
+    'SUF': 'green',
+    "SUF2": 'blue',
+    'DTSUF': 'purple'
 }
 
 COND_SHAPE = {
-    'no': 'r',
-    'view': 'y',
-    'yes': 'g'
+    'no': '-',
+    'view': '--',
+    'yes': ':'
+}
+
+COND_LEGEND = {
+    'no': '',
+    'view': '+view',
+    'yes': '+2x2'
 }
 
 def plot_benchmarks(runs, out_file='output.eps'):
     from matplotlib import pyplot as plt
+    from matplotlib.font_manager import FontProperties
+
+    fontP = FontProperties()
+    fontP.set_size('small')
 
     lines, densities = regroup_runs(runs)
 
     for line in lines:
         method = line['method']
-        compression = line['2x2']
+        condensation = line['2x2']
         ys = [line[x] for x in densities]
-        plt.plot(densities, ys, COND_SHAPE[compression] + METHOD_COLOR[method] + '-.')
-    plt.xlabel('Noise density')
-    plt.ylabel('ms')
-    plt.savefig(out_file, format='eps', dpi=1200)
+
+        plt.plot(densities, ys,
+                '.' + COND_SHAPE[condensation],
+                color=METHOD_COLOR[method],
+                label=method + COND_LEGEND[condensation])
+
+    plt.xlabel('Noise density, %')
+    plt.ylabel('CPU time, ms')
+
+    if '2000' not in out_file:
+        lgd = plt.legend(loc=2, prop=fontP)
+    else:
+        lgd = plt.legend(loc=2, bbox_to_anchor=(1, 1), prop=fontP)
+    plt.savefig(out_file,
+            bbox_extra_artists=(lgd,), bbox_inches='tight',
+            format='eps', dpi=1200)
 
 def write_benchmarks_latex(runs):
     headers = ['method', '2x2']

### 2016-02-08 utils/csv-to-latex.py
Mostly finish the paper, make the benchmark run 7h
Probably we should reorganize the benchmark:
add more tests of same size, streamline builds, etc.

@@ -62,6 +62,13 @@ METHOD_COLOR = {
     'DTSUF': 'purple'
 }
 
+METHOD_SHAPE = {
+    'DFS': '.',
+    'SUF': '.',
+    "SUF2": '.',
+    'DTSUF': '.'
+}
+
 COND_SHAPE = {
     'no': '-',
     'view': '--',
@@ -89,7 +96,7 @@ def plot_benchmarks(runs, out_file='output.eps'):
         ys = [line[x] for x in densities]
 
         plt.plot(densities, ys,
-                '.' + COND_SHAPE[condensation],
+                METHOD_SHAPE[method] + COND_SHAPE[condensation],
                 color=METHOD_COLOR[method],
                 label=method + COND_LEGEND[condensation])
 
@@ -104,30 +111,47 @@ def plot_benchmarks(runs, out_file='output.eps'):
             bbox_extra_artists=(lgd,), bbox_inches='tight',
             format='eps', dpi=1200)
 
-def write_benchmarks_latex(runs):
-    headers = ['method', '2x2']
-    grids = ['no', 'view', 'yes']
-    lines, densities = regroup_runs(runs)
 
-    columns = len(headers) + len(densities)
-    print(r'\begin{tabularx}{\textwidth}{| l  r  @{\extracolsep{\fill}}'
-            + 'r ' * (columns - 2) + '| }')
+def write_table(lines, densities):
+    HEADERS = ['method', '2x2']
+
+    def wrapsize(s):
+        #  return '\\scriptsize{' + str(s) + '}'
+        return str(s)
+
+    columns = len(HEADERS) + len(densities)
+
+    column_types = ['r'] * len(densities)
+    print(r'\begin{tabularx}{\linewidth}{| l  r  @{\extracolsep{\fill}}'
+            + ' '.join(column_types) + '| }')
 
     print(r'\cline{' + '{}-{}'.format(3, columns) + '}')
-    print(r'\multicolumn{' + str(len(headers)) + '}{c}{} & '
+    print(r'\multicolumn{' + str(len(HEADERS)) + '}{c}{} & '
         + r'\multicolumn{' + str(len(densities)) + '}{|c|}{density}', end='\\\\\n')
 
     print('\\hline')
-    print(' & '.join(headers + list(map(str, densities))), end='\\\\\n')
+    print(' & '.join(HEADERS + list(map(wrapsize, densities))), end='\\\\\n')
     print('\\hline')
-    for _, counter_lines in itertools.groupby(lines, lambda li: li[headers[0]]):
[…]

### 2016-02-08 paper/paper.tex

@@
 operation is $O(\alpha(HW))$, where $\alpha$ is the inverse Ackermann function,
-and $\alpha(HW) < 5$ for all remotely practical image sizes. % TODO links
+and $\alpha(HW) < 5$ for all remotely practical image sizes\cite{CLRS}. % TODO links
 
@@
 
-The idea is to run through each row twice: first time to establish label equality
+The idea is to run through each row twice: first time to establish label equivalence
 classes within row, and the second time to set unique equal label for all pixels of
@@
 
-The exact maximum number of label equality classes between two adjacent rows
+The exact maximum number of label equivalence classes between two adjacent rows
 depends on whether the given image is condensed.
-For non-condensed images $W=2M$, and maximum number of label equality classes
-in two adjacent rows is $M=\frac{W}2$, as two object pixels are both in adjacent
+For non-condensed images $W=2M$, and maximum number of label equivalence classes
+in two adjacent rows is $M=\frac{W}2$, as if two object pixels are both in adjacent
 rows and in adjacent columns, then they are neighbors in terms of 8-connectivity.
 
-On the other hand, for condensed images $W=M$, and maximum number of label equality
+On the other hand, for condensed images $W=M$, and maximum number of label equivalence
 class in two adjacent rows of cells is $2W=2M$.
@@
 This modification of SUF was implemented to compare the optimization we present with
-another optimization suggested by Wu, Otoo and Suzuki, % TODO reference
+another optimization suggested by Wu, Otoo and Suzuki\cite{wuotoo},
 namely decision tree.
@@
 each cell with 4 lookups to the original binary image.
-This is probably most interesting combined with SUF2 algorithm, as it is a way
-to win some time without increasing memory footprint.
+This is probably most interesting when combined with SUF2 algorithm, as it is a way
+to win some time while avoiding memory footprint increase.
 However this lazy approach increases the number of image lookups.
-So, according to our benchmark, this combination is efficient only for dense
-($>30\%$) noise.
+% So, according to our benchmark, this combination is efficient only for dense
+% ($>30\%$) noise.
 
@@
   \end{figure}
-  % \begin{table}[p]
-    % \centering
-    % \input{../bench-out/#1x#2}
-    % \caption{Average milliseconds of cpu time per run on noise images $#1 \times #2$ px}
-    % \label{tab:noise#1x#2}
-  % \end{table}
+  \begin{table}[p]
+    \centering
+    \input{tables/#1x#2}
+    \caption{Average CPU time per run on noise images $#1 \times #2$ px}
+    \label{tab:noise#1x#2}
+  \end{table}
 }
@@
 % \inputtable{200}{300}
-% \inputtable{400}{600}
+\inputtable{400}{600}
 \inputtable{800}{1200}
-% \inputtable{1200}{1800}
+\inputtable{1200}{1800}
 \inputtable{2000}{3000}
 \inputtable{4000}{6000}
-\inputtable{5000}{7500}
+% \inputtable{5000}{7500}
 
@@
 Методики выделения связных компонент в штриховых бинарных изображениях.
-Труды конференции ГрафиКон’2010. НИУ ИТМО. Санкт-Петербург. 2010, с. 169-174.
+Труды конференции ГрафиКон’2010. НИУ ИТМО. Санкт-Петербург. с. 169-174. (2010)
+
+\bibitem {CLRS}
+Cormen, H., Leiserson, E., Rivest, L., Stein, C.:
+Introduction to Algorithms (Second ed.), MIT Press (2001)
+
+\bibitem {wuotoo}
+Wu, K, Otoo, E., Suzuki, K.:
+Optimizing two-pass connected-component labeling algorithms.
+Pattern Anal. Appl. (2008)
 

### 2016-04-05 paper/paper.tex
Stage the reviewed state

@@
 \RequirePackage{amsmath}
-\documentclass{llncs}
+\documentclass[hidelinks]{llncs}
 
@@
 \renewcommand{\C}{\texttt{C} \xspace}
+\newcommand{\CXX}{\texttt{C++} \xspace}
 
@@
 \begin{abstract}
-Linear variation is a topological characteristic of a function of two variables.
-The problem of linear variation computing can be reduced to the problem
-of counting connected components in a binary image with eight-connected connectivity.
-The new hybrid method for the problem is presented.
-The method is essentially a modification of a two-scan algorithm for labeling
-that groups the pixels into $2 \times 2$ cells.
+The linear variation is a topological characteristic of the function of two variables.
+The problem of linear variation computing can be reduced to the problem of counting connected components in a binary image with eight-connected connectivity.
+%The new hybrid method for the problem is presented.
+%The method is essentially a modification of a two-scan algorithm for labeling
+%that groups the pixels into $2 \times 2$ cells.
+This is usually done through connected components labeling, and many approaches for that are known.
+This paper presents an efficient way to convert the initial binary image with 8-connectivity into a condensed (non-binary) image with new connectivity that is 2 times smaller in each dimension than the initial image.
+Most approaches for connected components labeling are still valid for the new image.
+As time and memory consumptions of conventional approaches usually depend linearly on the number of pixels,
+running them on the condensed image can be up to 4 times more efficient than running on the initial image.
+The method is essentially a modification of some known raster algorithms for labeling that groups pixels into $2 \times 2$ cells.
+A performance benchmark of the proposed  method applied to some approaches on noise images is provided.
 
-\keywords{Binary raster image, connected component labeling,
-pattern recognition}
+\keywords{Binary raster image, connected component labeling, pattern recognition}
 \end{abstract}
@@
 One of the most widely used approaches is total variation
-regularization [1]. % TODO fix bibliography
+regularization\cite{mak1}.
 Let $J(u)$ be the following functional:
@@
 characteristics referred to as linear variations.
-Kronrod [2] % TODO fix bibliography
+Kronrod\cite{mak2}
 introduced the notion of a regular component of the level set of a
@@
 one-dimensional tree of a function of two variables.
-In [3, 4, 5] % TODO fix bibliography
+In \cite{mak3,mak4,mak5}
 was discussed the necessity of the using a linear variation in the
@@
 This is usually done through CC labeling, and many approaches for that are
-known\cite{sterzh}.
+known\cite{hechao}.
 
@@
 
-So the benchmark tables show three levels of optimization applied to
+So the benchmark charts show three levels of optimization applied to
 DFS, SUF and SUF2:
-``no'' meaning no condensation, ``view'' meaning lazy memory-less condensation
-and ``yes'' meaning that a temporary condensed image was constructed in memory.
+just method name in legend refers to method applied to non-condensed image,
+``+view'' means lazy memory-less condensation
+and ``+2x2'' means that a temporary condensed image was constructed in memory.
 For the latter case, the construction time was included into the overall measured
@@
 
+\subsection{Measurement}
+
+For measurement purposes all algorithms and optimizations were implemented by us in \CXX.
+
+The program was compiled using \texttt{g++ 5.3.0-3} compiler with \texttt{-O3} optimization flag.
+All tests were run on same machine running under Arch Linux with standard \texttt{linux 4.4.1-2} kernel,
+utilising non-overclocked \texttt{Intel(R) Core(TM) i5-2430M CPU @ 2.40GHz} (L3 cache size 3072K) processor and non-overclocked 1333MHz 8GiB Kingston RAM.
+
+The test images were generated in the following manner. At first, 
+for given density $d$, the first $4dNM$ pixels of
+the $2N \times 2M$ image were filled with object values, while rest filled with background values.
+Then all the pixels were permuted using \texttt{std::shuffle} function from \CXX standard library,
+which is an implementation of Fisher-Yates Shuffle algorithm\cite{fisher:yates}.
+
+All methods were run on same tests, which was achieved by supplying same seed to the
+\texttt{std::mt19937} Mersenne Twister\cite{mt19937} random number engine implementation from \CXX standard library before passing the engine to \texttt{std::shuffle}.
+
 \subsection{Results}
 
-some results
+Line charts with benchmarks of mentioned algorithms on different noise images are shown on
+\Cref{fig:noise400x600,fig:noise800x1200,fig:noise1200x1800,fig:noise2000x3000,fig:noise4000x6000}.
+
+The extremely bigger CPU time for SUF and DTSUF methods on
+huge images could be explained by significant increase in the number of cache-misses
+produced by Union-Find.
+The charts prove effectiveness of the $2 \times 2$ condensation optimization.
 
@@
     \includegraphics[height=0.45\textheight, keepaspectratio]{pics/#1x#2.eps}
-    \caption{Average CPU time per run on noise images $#1 \times #2$ px}
+    \caption{Average CPU time per run on noise images $#1 \times #2$ px.}
     \label{fig:noise#1x#2}
   \end{figure}
-  \begin{table}[p]
-    \centering
-    \input{tables/#1x#2}
-    \caption{Average CPU time per run on noise images $#1 \times #2$ px}
-    \label{tab:noise#1x#2}
-  \end{table}
+ % \begin{table}[p]
+ %   \centering
+ %   \input{tables/#1x#2}
+ %   \caption{Average CPU time per run on noise images $#1 \times #2$ px}
+ %   \label{tab:noise#1x#2}
+ % \end{table}
 }
@@
 
-\section{Summary and future work}
+% \section{Summary and future work}
 
 \begin{thebibliography}{5}
+
+\bibitem {mak1}
+Rudin, L., Osher, S., Fatemi, E.: Nonlinear total variation based noised removal algorithms. Phys. D. 60, 259-268 (1992)
+
+\bibitem {mak2}
+Kronrod, A.: On functions of two variables ," Uspehi Mat. Nauk 5, 1(35), 24-134(1950)
+
+\bibitem {mak3}
+Makovetskii, A., Kober, V.: Image restoration based on topological properties of functions of two variables.  Proc. SPIE Applications of Digital Image Processing XXXV, 8499, 84990A (2012)
+
+\bibitem {mak4}
+4.	Makovetskii, A., Kober, V.: Modified gradient descent method for image restoration. Proc. SPIE Applications of Digital Image Processing XXXVI, 8856, 885608-1 (2013)
+
+\bibitem {mak5}
+Chochia, P., Milukova, O.: Comparison of Two-Dimensional Variations in the Context of the Digital Image Complexity Assessment.
+Journal of Communications Technology and Electronics, 2015, 60, no.
+12,  pp.
+1432-1440
 
@@
 
-\bibitem {sterzh}
-Стержанов, М.:
-Методики выделения связных компонент в штриховых бинарных изображениях.
-Труды конференции ГрафиКон’2010. НИУ ИТМО. Санкт-Петербург. с. 169-174. (2010)
+%\bibitem {sterzh}
+%Стержанов, М.:
+%Методики выделения связных компонент в штриховых бинарных изображениях.
+%Труды конференции ГрафиКон’2010. НИУ ИТМО. Санкт-Петербург. с. 169-174. (2010)
 
@@
 
+\bibitem {fisher:yates}
+ Fisher, A., Yates, F.:
+ Statistical tables for biological, agricultural and medical research (3rd ed.). 
+ London: Oliver \& Boyd. pp. 26–27. OCLC 14222135 (1948)
+
+\bibitem{mt19937}
+Matsumoto, M., Nishimura, T.:
+"Mersenne twister: a 623-dimensionally equidistributed uniform pseudo-random number generator".
+ACM Transactions on Modeling and Computer Simulation 8 (1): 3–30. (1998)
+
 \end{thebibliography}

### 2016-04-05 utils/csv-to-latex.py
Allow specifying image format and dpi

@@ -81,7 +81,7 @@ COND_LEGEND = {
     'yes': '+2x2'
 }
 
-def plot_benchmarks(runs, out_file='output.eps'):
+def plot_benchmarks(runs, out_file='output.eps', dpi=1200):
     from matplotlib import pyplot as plt
     from matplotlib.font_manager import FontProperties
 
@@ -109,7 +109,7 @@ def plot_benchmarks(runs, out_file='output.eps'):
         lgd = plt.legend(loc=2, bbox_to_anchor=(1, 1), prop=fontP)
     plt.savefig(out_file,
             bbox_extra_artists=(lgd,), bbox_inches='tight',
-            format='eps', dpi=1200)
+            format=out_file[-3:], dpi=dpi)
 
 
 def write_table(lines, densities):
@@ -156,6 +156,9 @@ def write_benchmarks_latex(runs):
 
 if __name__ == '__main__':
     runs = read_benchmarks_csv()
+    dpi = 1200
+    if '--dpi' in sys.argv:
+        dpi = int(sys.argv[sys.argv.index('--dpi')+1])
     if '--plot' in sys.argv:
         plot_benchmarks(runs, sys.argv[sys.argv.index('--plot')+1])
     write_benchmarks_latex(runs)

### 2016-04-05 utils/csv-to-latex.py
Enable specifying dpi for pictures

@@ -100,7 +100,7 @@ def plot_benchmarks(runs, out_file='output.eps', dpi=1200):
                 color=METHOD_COLOR[method],
                 label=method + COND_LEGEND[condensation])
 
-    plt.xlabel('Noise density, %')
+    plt.xlabel('density, %')
     plt.ylabel('CPU time, ms')
 
     if '2000' not in out_file:
@@ -160,6 +160,6 @@ if __name__ == '__main__':
     if '--dpi' in sys.argv:
         dpi = int(sys.argv[sys.argv.index('--dpi')+1])
     if '--plot' in sys.argv:
-        plot_benchmarks(runs, sys.argv[sys.argv.index('--plot')+1])
+        plot_benchmarks(runs, sys.argv[sys.argv.index('--plot')+1], dpi=dpi)
     write_benchmarks_latex(runs)
 

### 2016-04-06 paper/paper.tex
Add russian translation of key sections

@@
 \documentclass[hidelinks]{llncs}
+
+\usepackage[russian,english]{babel}
 
@@
 \usepackage{graphicx}
+
+% \newcommand{\texten}[1]{}
+% \newcommand{\textru}[1]{\foreignlanguage{russian}{\emph{#1}}{}}
+
+\newcommand{\texten}[1]{#1}
+\newcommand{\textru}[1]{\\
+  \foreignlanguage{russian}{\emph{#1}}}
 
@@
 
+\texten{
 We will use the fact that for the case of 8-connectivity all $4$ pixels of any
@@
 or ``cells''.
+}\textru{
+Наблюдение: с 8-связностью в любом квадратике $2 \times 2$ любые два пикселя
+находятся рядом. Значит если это пара объектных пикселей, то они находятся в одной
+КС, и значит каждый такой квадратик пересекается не более чем с одной КС.
+Поэтому вместо маркировки отдельных пикселей можно маркировать такие квадраты.
+Назовём их клетками.
+}
 
+\texten{
 Consider a $2N \times 2M$ binary image\footnote{
@@
   in both dimensions. If this is not the case, one can easily append a row or
-a column of background pixels to the image.}
+  a column of background pixels to the image.
+}
 specified by a binary predicate $b(x, y)$,
@@
 row index $y=0$, and the leftmost column has column index $x=0$.
+}\textru{
+Рассмотрим бинарное изображение размера $2N \times 2M$\footnote{
+  Для удобства мы рассматриваем только изображения с чётной высотой и шириной.
+  В общем случае при необходимости можно добавить строку или столбец с фоновыми
+  пикселями.
+},
+заданное предикатом $b(x, y)$, который возвращает $1$, если переданы координаты
+объектного пикселя, и $0$ иначе.
+Мы везде будем использовать $0$-индексацию: верхняя строка и левый столбец имеют
+номер $0$.
+}
 
+\texten{
 The condensed image will be of size $N \times M$.
@@
 to be the following:
+}\textru{
+Сжатое изображение будет иметь размер $N \times M$.
+Каждый пиксель полученного изображения должен будет однозначно задавать конфигурацию
+соответствующих четырёх пикселей исходного изображения, поэтому он может быть одного
+из $2^4 = 16$ цветов. Предлагается такая функция $c(x,y)$, кодирующая четыре исходных
+пикселя в цвет клетки:
+}
 \begin{align*}
@@
 
+\texten{
 So the value of each pixel of the initial image is stored in the corresponding
@@
 pixels inside one cell is shown on \autoref{fig:pixelsEnumeration}.
+}\textru{
+Таким образом значение каждого из исходных пикселей просто записано в соответствующий
+бит в итоговом пикселе. Важно отметить, что если все четыре исходных пикселя были
+фоновые, то и клетка будет иметь цвет 0. Нумерация пикселей внутри клетки показана на
+\autoref{fig:pixelsEnumeration}.
+}
 
@@
 
+\texten{
 We need also to define connectivity function for the condensed image.
@@
 configuration. See \autoref{fig:connectivity:examples} for examples.
+}\textru{
+Ещё нужно дать определение связности для сжатого изображения. Мы уже не можем
+использовать 8-связность: теперь нельзя сказать, связаны ли две клетки, не
+проанализировав их цвета. Смотри \autoref{fig:connectivity:examples}.
+}
 
@@
 
+\texten{
 However, for every variant of relative placement of two nearby cells there is
@@
 The masks are shown in \autoref{fig:connectivity:masks}.
+}\textru{
+Тем не менее, можно заметить, что анализировать можно не все биты цвета, а только
+некоторую маску тех бит, которые важны для связности двух конкретных клеток.
+Эти маски показаны на \autoref{fig:connectivity:masks}.
+}
 
@@
 
+\texten{
 The idea is that if we naturally encode mask pixels as 4-bit numbers in the
@@
 factor of 4.
+}\textru{
+Идея в том, что с нашим естественным способом кодирования конфигурации клеток в
+4-разрядных числах, предикат, определяющий связность двух соседных клеток,
+может быть представлен всего двумя битовыми и одной логической операцией И.
+В качестве примера, на \autoref{fig:connectivity:code} приведено сравнение
+возможных фрагментов программы на языке \C\hspace{-0.8em}, проверяющих, связан
+ли текущий пиксель (или клетка) с пикселем (или клеткой) над ним (ней).
+Стоит отметить, что хотя конструкция
+усложнилась, битовые операции как правило очень эффективны, и это усложнение, скорее
+всего, незначительно по сравнению с уменьшением количества рассматриваемых позиций в 4
+раза.
+}
 
@@
 
+\texten{
 As each of these methods, omitting DTSUF could be used with both condensed and
@@
 on input of each algorithm as $H$ and $W$.
+}\textru{
+Так как все рассматриваемые методы, кроме DTSUF, можно применять как для сжатого,
+так и для несжатого изображения, мы обозначим размеры входного изображения
+каждого изображения за $H$ и $W$.
+}
 
+\texten{
 Also to avoid repeating ``pixel or cell'' we will just use the word ``pixel''
@@
 to it in terms of appropriate connectivity.
+}\textru{
+По аналогии, чтобы избежать повторения ``пиксель или клетка'' мы будет просто
+использовать слово ``пиксель'', имея в виду растр входного изображения.
+Под соседями пикселя будем понимать пиксели, прямо связанные с ним в терминах
+соответствующей связности.
+}
 
@@
 
+\texten{
 Another general approach for graph CC labeling which is more popular in image
@@
 that are considered to be in the same CC.
+}\textru{
+Другой, более популярный в этом контексте общий подход к подсчёту компонент
+связности состоит в том, чтобы использовать структуру данных для поддержания
+набора непересекающихся множеств пикселей, чем, по сути, КС и являются.
+}
 
+\texten{
 The structure is commonly called Union-Find, as the two essential operations
@@
 and $\alpha(HW) < 5$ for all remotely practical image sizes\cite{CLRS}. % TODO links
+}\textru{
+Эту структуру принято называть Union-Find, так как две основные операции,
+которые она поддерживает, это Union~--- объединение двух множеств в одно, и
+Find~--- нахождение характерного представителя для множества, содержащего данный
+элемент.
+Этот представитель может быть использован для проверки принадлежности двух
+элементов одному и тому же множеству.
+Эта структура данных хорошо изучена и может быть реализована так, что
+амортизированная сложность каждой операции~--- всего $O(\alpha(HW))$, где
+$\alpha$~--- обратная функция Аккермана.
+Для всех практически возможных размеров изображения эта функция не превосходит 5.
+}
 
-The approach that is named SUF (Scan plus Union-Find) in this paper follows.
+\texten{
+The approach that is named SUF (Scan with Union-Find) in this paper follows.
+}\textru{
+Далее следует описание подхода, который мы называем SUF (Scan with Union-Find).
+}
 
+\texten{
 Pixels are processed in the natural order: from top to bottom and from
@@
 that were already processed, see \autoref{fig:suf:neighbors}.
+}\textru{
+Пиксели обрабатываются в естественном порядке сверху вниз слева направо, так что
+каждая строка начинает обрабатываться, когда все предыдущие строки уже
+обработаны. Таким образом, у очередного пикселя может быть до четырёх пикселей,
+которые находятся рядом и уже были обработаны.
+}
 
+\texten{
 At first each object pixel is considered to belong to its own CC, so right before
@@
 the set number $xW + y$ to guarantee absence of collisions.
+}\textru{
+Каждый объектный пиксель сначала предполагается принадлежащим своей КС.
+Поэтому прежде чем начинать рассматривать соседей, мы увеличиваем счётчик.
+Более того того, каждый пиксель изначально ассоциирован со своим собственным
+независимым множеством в структуре Union-Find, состоящем из одного элемента.
+Конкретно пиксель в позиции $x, y$ заранее ассоциирован с элементом номер $xW + y$,
+чтобы гарантировать отсутствие коллизий.
+}
 
+\texten{
 The possible neighbors are processed in the order shown on \autoref{fig:suf:neighbors},
@@
 CC, the CC counter is decremented.
+}\textru{
+Возможные соседи перебираются в порядке, показанном на
+\autoref{fig:suf:neighbors}.
+Для каждого объектного пикселя, который напрямую соседствует с текущим, мы
+применяем операцию Union для множества, содержащих текущий пиксель и
+множества, содержащего этого соседа.
+Если эти множества не совпадали, то мы нашли место, где КС соединяются.
+В этом случае уменьшаем счётчик на 1.
+}
 
@@
 
+\texten{
 Note that this algorithm outputs not the matrix of final labels, but only the number
@@
 The latter is exactly what is needed for the aforementioned linear variation problem.
+}\textru{
+Отметим, что этот алгоритм выдаёт не матрицу финальных меток, а только
+количество компонент связности.
+Именно в этом месте он отличается от некоторых известных подходов к CC labeling.
+Для текущей задачи нам нужно именно считать компоненты связности.
+}
 
+\texten{
 Also we have to mention that for performance reasons the Union-Find internal structures
@@
 coordinates. So memory consumption of SUF is linear in the number of pixels.
+}\textru{
+Также необходимо отметить, что для лучшей производительности служебные массивы
+структуры Union-Find заранее выделяются размера как минимум $HW$, чтобы
+обеспечить простоту соответствия между пикселями изображения и элементами в
+системе множеств.
+Таким образом, SUF использует линейное от числа пикселей количество памяти
+$O(HW)$.
+}
 
@@
 
+\texten{
 This approach is a modification of SUF designed to reduce memory consumption
 and potential number of cache-misses generated by Union-Find operations.
+}\textru{
+Этот подход~--- модификация SUF, предложенная, чтобы сократить потребление
+памяти и предположительно сильно уменьшить число кэш-промахов, происходящих
+из-за произвольного доступа к большим массивам структуры Union-Find.
+}
 
+\texten{
 The idea is to run through each row twice: first time to establish label equivalence
@@
 for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.
+}\textru{
+Идея в том, чтобы бежать по каждой строке дважды: один раз чтобы установить
+классы эквивалентности меток в текущей и предыдущей строках, и второй раз, чтобы
+расставить уникальные метки в пиксели каждого класса.
+Это даёт нам возможность иметь не более $O(W)$ различных меток в каждый момент
+времени, и поэтому Union-Find можно выделять размера всего $O(W)$, при том что в
+SUF выделялось ровно $WH$ элементов.
+}
 
+\texten{
 The exact maximum number of label equivalence classes between two adjacent rows
@@
 rows and in adjacent columns, then they are neighbors in terms of 8-connectivity.
+}\textru{
+Точная оценка на максимальное число классов эквивалентности меток в двух
+соседних строках зависит от того, сжатое изображение поступило на вход, или нет.
+Для несжатых изображений $W=2M$, и наибольшее число классов эквивалентности в
+соседних строках равно $M=\frac{W}2$, так как если два объектных пикселя
+находятся одновременно в соседних строках и соседних столбцах, то они соседи в
+смысле 8-связности.
+}
 
+\texten{
 On the other hand, for condensed images $W=M$, and maximum number of label equivalence
@@
 See \autoref{fig:suf2:extremal} for illustration on the extremal cases.
+}\textru{
+С другой стороны, для сжатых изображений $W=M$, и максимальное число классов
+эквивалентности равно $2W=2M$. Смотри примеры на \autoref{fig:suf2:extremal}.
+}
 
@@
 
+\texten{
 Our benchmarks show that, although SUF2 without condensation is generally more
@@
 small and medium-sized noise images.
+}\textru{
+Проведённые эксперименты показывают, что хотя SUF2 без сжатия в целом более
+эффективен, чем SUF без сжатия, приведённая оценка делает относительное
+ускорение SUF2 со сжатием не таким значительным.
+}
 
+\texten{
 Note that, while SUF2 is intended for CC counting, its memory
@@
 other resource-tight applications.
+}\textru{
+Отметим, что SUF2 решает задачу подсчёта СК, используя всего $O(W)$
+дополнительной памяти.
+}
 
@@
 
+\texten{
 This modification of SUF was implemented to compare the optimization we present with
@@
 applicable for condensed image, as it exploits similar observations on 8-connectivity.
+}\textru{
+Эта модификация SUF была добавлена чтобы сравнить нашу оптимизацию с
+одной из оптимизаций, предложенных Wu, Otoo и Suzuki\cite{wuotoo}, а именно с
+деревом принятия решения.
+К сожалению, DTSUF (Decision Tree Scan with Union-Find) не применим для сжатого
+изображения, так как использует свойства 8-связности, которыми наша новая
+связность не обладает.
+}
 
+\texten{
 As before, the goal is to reduce the number of pixel lookup and label copy operations.
@@
 shown on \autoref{fig:dtsuf}.
+}\textru{
+Как и раньше, цель состоит в том, чтобы уменьшить число запросов на содержание
+пикселя и копирований меток.
+Это достигается аккуратным рассмотрением соседей в оптимальном порядке и
+остановкой рассмотрения, как только достаточная для выполнения всех необоходимых
+действий информация была получена.
+Смотри дерево на \autoref{fig:dtsuf}.
+}
 

### 2016-04-14 paper/paper.tex
Fix references again

@@
 
-% \newcommand{\texten}[1]{}
-% \newcommand{\textru}[1]{\foreignlanguage{russian}{\emph{#1}}{}}
+\newcommand{\texten}[1]{#1}
+\newcommand{\textru}[1]{}
 
-\newcommand{\texten}[1]{#1}
-\newcommand{\textru}[1]{\\
-  \foreignlanguage{russian}{\emph{#1}}}
+%% Чтобы переключиться на русский язык в некоторых секциях, можно 
+%% раскомментировать следующие две команды
+% \renewcommand{\texten}[1]{}
+% \renewcommand{\textru}[1]{\foreignlanguage{russian}{\emph{#1}}{}}
+
+%% Чтобы оставить оба языка, можно раскомментировать
+%% следующие две команды
+% \renewcommand{\texten}[1]{#1}
+% \renewcommand{\textru}[1]{\\ \foreignlanguage{russian}{\emph{#1}}}
 
@@
 characteristics referred to as linear variations.
-Kronrod\cite{mak2}
+Kronrod%\cite{mak2}
 introduced the notion of a regular component of the level set of a
@@
 
+\texten{
 For memory-tight applications such as embedded systems it can be useful to
@@
 % ($>30\%$) noise.
+}\textru{
+В некоторых случаях может быть более приемлемо использовать сжатие неявно, то
+есть вместо построения 
+}
 
@@
 
-The test images were generated in the following manner. At first, 
-for given density $d$, the first $4dNM$ pixels of
-the $2N \times 2M$ image were filled with object values, while rest filled with background values.
-Then all the pixels were permuted using \texttt{std::shuffle} function from \CXX standard library,
-which is an implementation of Fisher-Yates Shuffle algorithm\cite{fisher:yates}.
+The test images were generated in the following manner. At first, for given
+density $d$, the first $4dNM$ pixels of the $2N \times 2M$ image were filled
+with object values, while rest filled with background values. Then all the
+pixels were permuted using \texttt{std::shuffle} function from \CXX standard
+library, which is an implementation of Fisher-Yates Shuffle
+algorithm%\cite{fisher:yates}
+.
 
@@
 
-Line charts with benchmarks of mentioned algorithms on different noise images are shown on
-\Cref{fig:noise400x600,fig:noise800x1200,fig:noise1200x1800,fig:noise2000x3000,fig:noise4000x6000}.
+Line charts with benchmarks of mentioned algorithms on different noise images
+are shown on
+\Cref{fig:overall:times}.
 
-The extremely bigger CPU time for SUF and DTSUF methods on
-huge images could be explained by significant increase in the number of cache-misses
-produced by Union-Find.
-The charts prove effectiveness of the $2 \times 2$ condensation optimization.
+The extremely bigger CPU time for SUF and DTSUF methods on huge images could be
+explained by significant increase in the number of cache-misses produced by
+Union-Find. The charts prove effectiveness of the $2 \times 2$ condensation
+optimization.
 
-\newcommand{\inputtable}[2]{
 \begin{figure}
   \centering
-    \includegraphics[height=0.45\textheight, keepaspectratio]{pics/#1x#2.eps}
-    \caption{Average CPU time per run on noise images $#1 \times #2$ px.}
-    \label{fig:noise#1x#2}
+  \includegraphics[height=0.45\textheight, keepaspectratio]{pics/overall.eps}
+  \caption{Average CPU time per run on random images $1200 \times 1800$ px.}
+  \label{fig:overall:times}
 \end{figure}
- % \begin{table}[p]
- %   \centering
- %   \input{tables/#1x#2}
- %   \caption{Average CPU time per run on noise images $#1 \times #2$ px}
- %   \label{tab:noise#1x#2}
- % \end{table}
+
+\newcommand{\inputnorm}[1]{
+  \begin{subfigure}[t]{0.45\linewidth}
+    \centering
+    \includegraphics[width=\textwidth,
+                     keepaspectratio]{pics/normed_#1.eps}
+    \caption{Some.}
+    \label{fig:normed:#1}
+  \end{subfigure}
 }
 
-% \inputtable{200}{300}
-\inputtable{400}{600}
-\inputtable{800}{1200}
-\inputtable{1200}{1800}
-\inputtable{2000}{3000}
-\inputtable{4000}{6000}
-% \inputtable{5000}{7500}
+\begin{figure}
+  \centering
+  \inputnorm{DFS}
+  \inputnorm{SUF}
+  \inputnorm{SUF2}
+  \inputnorm{DTSUF}
+  \label{fig:normed}
+\end{figure}
 
@@
 \bibitem {mak1}
-Rudin, L., Osher, S., Fatemi, E.: Nonlinear total variation based noised removal algorithms. Phys. D. 60, 259-268 (1992)
+Rudin, L., Osher, S., Fatemi, E.:
+Nonlinear total variation based noised removal algorithms.
+Phys. D. 60, 259-268 (1992)
 
-\bibitem {mak2}
-Kronrod, A.: On functions of two variables ," Uspehi Mat. Nauk 5, 1(35), 24-134(1950)
+% \bibitem {mak2}
+% Kronrod, A.:
+% On functions of two variables.
+% Uspehi Mat. Nauk 5, 1(35), 24-134(1950)
 
 \bibitem {mak3}
-Makovetskii, A., Kober, V.: Image restoration based on topological properties of functions of two variables.  Proc. SPIE Applications of Digital Image Processing XXXV, 8499, 84990A (2012)
+Makovetskii, A., Kober, V.:
+Image restoration based on topological properties of functions of two variables.
+Proc. SPIE Applications of Digital Image Processing XXXV, 8499, 84990A (2012)
 
 \bibitem {mak4}
-4.	Makovetskii, A., Kober, V.: Modified gradient descent method for image restoration. Proc. SPIE Applications of Digital Image Processing XXXVI, 8856, 885608-1 (2013)
+Makovetskii, A., Kober, V.:
+Modified gradient descent method for image restoration.
+Proc. SPIE Applications of Digital Image Processing XXXVI, 8856, 885608-1 (2013)
 
 \bibitem {mak5}
-Chochia, P., Milukova, O.: Comparison of Two-Dimensional Variations in the Context of the Digital Image Complexity Assessment.
-Journal of Communications Technology and Electronics, 2015, 60, no.
-12,  pp.
-1432-1440
+Chochia, P., Milukova, O.:
+Comparison of Two-Dimensional Variations in the Context of the Digital Image
+Complexity Assessment.
+Journal of Communications Technology and Electronics, 2015, 60, no. 12,  pp. 1432-1440
 
@@
 He, L., Chao, Y., Suzuki, K., Wu, K.:
-Fast connected-component labeling. Pattern Recognition. v42, Pages 1977-1987. (2009)
+Fast connected-component labeling.
+Pattern Recognition. v42, Pages 1977-1987. (2009)
 
@@
 Cormen, H., Leiserson, E., Rivest, L., Stein, C.:
-Introduction to Algorithms (Second ed.), MIT Press (2001)
+Introduction to Algorithms (Second ed.).
+MIT Press (2001)
 
@@
 
-\bibitem {fisher:yates}
- Fisher, A., Yates, F.:
- Statistical tables for biological, agricultural and medical research (3rd ed.). 
- London: Oliver \& Boyd. pp. 26–27. OCLC 14222135 (1948)
+% \bibitem {fisher:yates}
+% Fisher, A., Yates, F.:
+% Statistical tables for biological, agricultural and medical research (3rd ed.).
+% London: Oliver \& Boyd. pp. 26–27. OCLC 14222135 (1948)
 
@@
 Matsumoto, M., Nishimura, T.:
-"Mersenne twister: a 623-dimensionally equidistributed uniform pseudo-random number generator".
+Mersenne twister: a 623-dimensionally equidistributed uniform pseudo-random
+number generator.
 ACM Transactions on Modeling and Computer Simulation 8 (1): 3–30. (1998)

### 2016-04-14 paper/errata.tex
Fix references

\documentclass{article}

\usepackage{cmap}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}

\usepackage{xr}
\externaldocument{paper}

\begin{document}

\title{Errata for Linear Variation and an Optimization of Algorithms for Connected
Components Labeling in a Binary Image}

\author{Fedor Alekseev, Mikhail Alekseev and Artyom Makovetskii}

\date{}

\maketitle

\begin{enumerate}
  \item The references page was revised: references on Kronrod and Fisher-Yates
    were removed, as well the extra number before one of the references on
    Makovetskii and Kober.
\end{enumerate}

\end{document}

### 2016-04-14 paper/paper.tex
Fix references

--- 2016-04-05 paper/paper.tex
@@
 
-\institute{Moscow Institute of Physics and Technology State University, Dolgoprudny, Russia\\
+\institute{Moscow Institute of Physics and Technology, Dolgoprudny, Russia\\
 \email{alekseev@phystech.edu}
@@
 \begin{abstract}
-The linear variation is a topological characteristic of the function of two variables.
-The problem of linear variation computing can be reduced to the problem of counting connected components in a binary image with eight-connected connectivity.
-%The new hybrid method for the problem is presented.
-%The method is essentially a modification of a two-scan algorithm for labeling
-%that groups the pixels into $2 \times 2$ cells.
-This is usually done through connected components labeling, and many approaches for that are known.
-This paper presents an efficient way to convert the initial binary image with 8-connectivity into a condensed (non-binary) image with new connectivity that is 2 times smaller in each dimension than the initial image.
-Most approaches for connected components labeling are still valid for the new image.
-As time and memory consumptions of conventional approaches usually depend linearly on the number of pixels,
-running them on the condensed image can be up to 4 times more efficient than running on the initial image.
-The method is essentially a modification of some known raster algorithms for labeling that groups pixels into $2 \times 2$ cells.
-A performance benchmark of the proposed  method applied to some approaches on noise images is provided.
+The linear variation is a topological characteristic of the function of two
+variables. The problem of linear variation computing can be reduced to the
+problem of counting connected components in a binary image with eight-connected
+connectivity.
+%The new hybrid method for the problem is presented. The method is essentially
+%a modification of a two-scan algorithm for labeling that groups the pixels
+%into $2 \times 2$ cells.
+This is usually done through connected components labeling, and many approaches
+for that are known. This paper presents an efficient way to convert the initial
+binary image with 8-connectivity into a condensed (non-binary) image with new
+connectivity that is 2 times smaller in each dimension than the initial image.
+Most approaches for connected components labeling are still valid for the new
+image. As time and memory consumptions of conventional approaches usually
+depend linearly on the number of pixels, running them on the condensed image
+can be up to 4 times more efficient than running on the initial image. The
+method is essentially a modification of some known raster algorithms for
+labeling that groups pixels into $2 \times 2$ cells. A performance benchmark of
+the proposed  method applied to some approaches on noise images is provided.
 
@@
 
-The norm and modulus of the gradient are metrical characteristic of a function of
-two variables.
+The norm and modulus of the gradient are metrical characteristic of a function
+of two variables.
 Continuous functions of two variables also have a set of topological
 characteristics referred to as linear variations.
-Kronrod\cite{mak2}
+Kronrod%\cite{mak2}
 introduced the notion of a regular component of the level set of a
 continuous function of two variables.
-The simplest topological characteristic in the linear variation theory is a number
-of regular components for all level sets of a function.
+The simplest topological characteristic in the linear variation theory is a
+number of regular components for all level sets of a function.
 Full information about linear variations of a function is contained in the
@@
 
-Let $w$ be a binary discrete function $w=(w_{i,j})$, where $w_{i,j} \in \{0,1\}$,
-for all pairs $(i,j)$.
-A subset of such pairs $(i, j)$ when $w_{i,j}=1$ and all elements of the subset are
-connected by the 8-connectivity, is called the connected component of the binary
-function $w$.
-For a number $k \in \N$ and a discrete function $u$ we define the following
-indicator function $\chi$:
+Let $w$ be a binary discrete function $w=(w_{i,j})$, where $w_{i,j} \in
+\{0,1\}$, for all pairs $(i,j)$. A subset of such pairs $(i, j)$ when
+$w_{i,j}=1$ and all elements of the subset are connected by the 8-connectivity,
+is called the connected component of the binary function $w$. For a number $k
+\in \N$ and a discrete function $u$ we define the following indicator function
+$\chi$:
 \begin{equation}
@@
 \begin{definition}
-  The number $V_k(u_{i,j})$ of connected components for a level $k$, $k \in \N$ of
-  the discrete function $u$ is called the number of connected components of the 
-  binary discrete function $\chi_k(u_{i,j})$.
+  The number $V_k(u_{i,j})$ of connected components for a level $k$, $k \in \N$
+  of the discrete function $u$ is called the number of connected components of
+  the binary discrete function $\chi_k(u_{i,j})$.
 \end{definition}
@@
 
-This paper presents an efficient way to convert the initial binary image
-with 8-connectivity into
-a condensed (non-binary) image with new connectivity that is $2$ times smaller in
-each dimension than the initial image.
-Most approaches for CC labeling are still valid for the new image.
-As time and memory consumptions of conventional approaches usually depend linearly on
-the number of pixels, running them on the condensed image can be
-more efficient than running on the initial image.
+This paper presents an efficient way to convert the initial binary image with
+8-connectivity into a condensed (non-binary) image with new connectivity that
+is $2$ times smaller in each dimension than the initial image. Most approaches
+for CC labeling are still valid for the new image. As time and memory
+consumptions of conventional approaches usually depend linearly on the number
+of pixels, running them on the condensed image can be more efficient than
+running on the initial image.
 
-It is worth noting that our method exploits an important property of 8-connectivity
-that does not hold for 4-connectivity, so it is not applicable for the latter case.
+It is worth noting that our method exploits an important property of
+8-connectivity that does not hold for 4-connectivity, so it is not applicable
+for the latter case.
 
-\section{$2\times2$ condensation}
+\section{$2 \times 2$ condensation}
 
 We will use the fact that for the case of 8-connectivity all $4$ pixels of any
-$2 \times 2$ square are adjacent to each other, and so are contained in the same CC
-and should have same labels upon completion of the labeling algorithm.
-So instead of labeling single pixels we can label $2 \times 2$ groups of pixels,
-or ``cells''.
+$2 \times 2$ square are adjacent to each other, and so are contained in the
+same CC and should have same labels upon completion of the labeling algorithm.
+So instead of labeling single pixels we can label $2 \times 2$ groups of
+pixels, or ``cells''.
 
@@
 specified by a binary predicate $b(x, y)$,
-returning $1$ in case the pixel at the given coordinates is black (object pixel),
-and $0$ otherwise (background pixel).
-For convenience we use 0-indexation throughout the paper, so the topmost row has
-row index $y=0$, and the leftmost column has column index $x=0$.
+returning $1$ in case the pixel at the given coordinates is black (object
+pixel), and $0$ otherwise (background pixel). For convenience we use
+0-indexation throughout the paper, so the topmost row has row index $y=0$, and
+the leftmost column has column index $x=0$.
 
@@
 Each pixel of the condensed image will unambiguously represent the configuration
-of the corresponding $2\times 2$ cell of pixels of the initial binary image.
+of the corresponding $2 \times 2$ cell of pixels of the initial binary image.
 So the condensed image will contain pixels of $2^4 = 16$ colors.
-We suggest the function $c(x, y)$ denoting the color of a pixel of condensed image
-to be the following:
+We suggest the function $c(x, y)$ denoting the color of a pixel of condensed
+image to be the following:
 \begin{align*}
@@
 bit in color of the corresponding pixel of the condensed image.
-Note that if $c(x, y) = 0$ for some $(x, y)$, then there are no object pixels in
-this cell, and no label should be assigned to this whole cell. The enumeration of
-pixels inside one cell is shown on \autoref{fig:pixelsEnumeration}.
+Note that if $c(x, y) = 0$ for some $(x, y)$, then there are no object pixels
+in this cell, and no label should be assigned to this whole cell. The
+enumeration of pixels inside one cell is shown on
+\autoref{fig:pixelsEnumeration}.
 
@@
 in terms of 8-connectivity and each of them contains at least one object pixel
-in the mask of important pixels corresponding to relative placement of the cells.
-The masks are shown in \autoref{fig:connectivity:masks}.
+in the mask of important pixels corresponding to relative placement of the
+cells. The masks are shown in \autoref{fig:connectivity:masks}.
 
@@
 operations.
-As an example, \autoref{fig:connectivity:code} compares possible codes in \C checking
-if the labels of the current pixel or cell and the one right on top of it
-should be same for classic 8-connectivity and for the new connectivity.
-Note that as bitwise operations are usually relatively cheap, we added only a
-little complication compared to decreasing the number of pixels to process by the 
-factor of 4.
+As an example, \autoref{fig:connectivity:code} compares possible codes in \C
+checking if the labels of the current pixel or cell and the one right on top of
+it should be same for classic $8$-connectivity and for the new connectivity. Note
+that as bitwise operations are usually relatively cheap, we added only a little
+complication compared to decreasing the number of pixels to process by the
+factor of $4$.
 
@@
 meaning the appropriate raster for non-condensed and condensed images.
-When we mention neighbors of some pixel, we mean pixels that are directly connected
-to it in terms of appropriate connectivity.
+When we mention neighbors of some pixel, we mean pixels that are directly
+connected to it in terms of appropriate connectivity.
 
@@
 
-The first algorithm we chose for the benchmark is Depth First Search graph traversal.
-Graph traversal algorithms are the general approach for CC labeling in graphs, and
-DFS is arguably the most simple.
-Although not intended for image labeling and
-suboptimal in most cases, it still has linear in the number of pixels worst-case
-complexity in both time and memory.
+The first algorithm we chose for the benchmark is Depth First Search graph
+traversal. Graph traversal algorithms are the general approach for CC labeling
+in graphs, and DFS is arguably the most simple. Although not intended for image
+labeling and suboptimal in most cases, it still has linear in the number of
+pixels worst-case complexity in both time and memory.
 
-Like most approaches, the algorithm has an outer loop over all pixels of the image.
-Each time an object pixel (non-empty cell, in case it is actually a cell)
-which is not labeled yet is encountered, DFS traversal is launched from that pixel.
-The traversal labels all the object pixels of the same CC, launching recursively
-from all unlabeled neighbors of current pixel.
+Like most approaches, the algorithm has an outer loop over all pixels of the
+image. Each time an object pixel (non-empty cell, in case it is actually a
+cell) which is not labeled yet is encountered, DFS traversal is launched from
+that pixel. The traversal labels all the object pixels of the same CC,
+launching recursively from all unlabeled neighbors of current pixel.
 
@@
 Another general approach for graph CC labeling which is more popular in image
-labeling is using some kind of a data structure to maintain disjoint sets of pixels
-that are considered to be in the same CC.
+labeling is using some kind of a data structure to maintain disjoint sets of
+pixels that are considered to be in the same CC.
 
-The structure is commonly called Union-Find, as the two essential operations
-it is supposed to support are Union, which is used to merge some two of the
-disjoint sets in one, and Find, which is used to find the representative element
-of the set some element belongs to, so that we can always tell if two elements are
-in the same set.
-The structure can be implemented so that the amortized time complexity for each
-operation is $O(\alpha(HW))$, where $\alpha$ is the inverse Ackermann function,
-and $\alpha(HW) < 5$ for all remotely practical image sizes\cite{CLRS}. % TODO links
+The structure is commonly called Union-Find, as the two essential operations it
+is supposed to support are Union, which is used to merge some two of the
+disjoint sets in one, and Find, which is used to find the representative
+element of the set some element belongs to, so that we can always tell if two
+elements are in the same set. The structure can be implemented so that the
+amortized time complexity for each operation is $O(\alpha(HW))$, where $\alpha$
+is the inverse Ackermann function, and $\alpha(HW) < 5$ for all remotely
+practical image sizes\cite{CLRS}. % TODO links
 
@@
 
-At first each object pixel is considered to belong to its own CC, so right before
-considering the neighbors we increment the CC counter.
-Moreover, it is associated with its own disjoint set in the Union-Find structure.
-More specifically, object pixel at position $x, y$ is associated with
-the set number $xW + y$ to guarantee absence of collisions.
+At first each object pixel is considered to belong to its own CC, so right
+before considering the neighbors we increment the CC counter. Moreover, it is
+associated with its own disjoint set in the Union-Find structure. More
+specifically, object pixel at position $x, y$ is associated with the set number
+$xW + y$ to guarantee absence of collisions.
 
-The possible neighbors are processed in the order shown on \autoref{fig:suf:neighbors},
-and for each neighbor that is object pixel, the Union operation is performed
-on the two sets containing the current and the neighbor pixel.
-In case the Union operation was successful, which means that the pixels
-had belonged previously to different sets and we have just found a way to merge two
-CC, the CC counter is decremented.
+The possible neighbors are processed in the order shown on
+\autoref{fig:suf:neighbors}, and for each neighbor that is object pixel, the
+Union operation is performed on the two sets containing the current and the
+neighbor pixel. In case the Union operation was successful, which means that
+the pixels had belonged previously to different sets and we have just found a
+way to merge two CC, the CC counter is decremented.
 
@@
 
-Note that this algorithm outputs not the matrix of final labels, but only the number
-of CC.
-This is the place in which it differs from some common two-pass approaches for CC % TODO references
-labeling, as it is an algorithm not for labeling, but just for counting CC.
-The latter is exactly what is needed for the aforementioned linear variation problem.
+Note that this algorithm outputs not the matrix of final labels, but only the
+number of CC. This is the place in which it differs from some common two-pass
+approaches for CC % TODO references
+labeling, as it is an algorithm not for labeling, but just for counting CC. The
+latter is exactly what is needed for the aforementioned linear variation
+problem.
 
-Also we have to mention that for performance reasons the Union-Find internal structures
-are allocated ahead to have the size of at least $HW$ to guarantee that
-each object can be easily associated with unique starting set given only its
-coordinates. So memory consumption of SUF is linear in the number of pixels.
+Also we have to mention that for performance reasons the Union-Find internal
+structures are allocated ahead to have the size of at least $HW$ to guarantee
+that each object can be easily associated with unique starting set given only
+its coordinates. So memory consumption of SUF is linear in the number of
+pixels.
 
@@
 
-The idea is to run through each row twice: first time to establish label equivalence
-classes within row, and the second time to set unique equal label for all pixels of
-same CC.
-This gives us the opportunity to maintain just $O(W)$
-different labels at each moment, and so the Union-Find structure can be allocated
-for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.
+The idea is to run through each row twice: first time to establish label
+equivalence classes within row, and the second time to set unique equal label
+for all pixels of same CC. This gives us the opportunity to maintain just
+$O(W)$ different labels at each moment, and so the Union-Find structure can be
+allocated for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.
 
 The exact maximum number of label equivalence classes between two adjacent rows
-depends on whether the given image is condensed.
-For non-condensed images $W=2M$, and maximum number of label equivalence classes
-in two adjacent rows is $M=\frac{W}2$, as if two object pixels are both in adjacent
-rows and in adjacent columns, then they are neighbors in terms of 8-connectivity.
+depends on whether the given image is condensed. For non-condensed images
+$W=2M$, and maximum number of label equivalence classes in two adjacent rows is
+$M=\frac{W}2$, as if two object pixels are both in adjacent rows and in
+adjacent columns, then they are neighbors in terms of 8-connectivity.
 
-On the other hand, for condensed images $W=M$, and maximum number of label equivalence
-class in two adjacent rows of cells is $2W=2M$.
-See \autoref{fig:suf2:extremal} for illustration on the extremal cases.
+On the other hand, for condensed images $W=M$, and maximum number of label
+equivalence class in two adjacent rows of cells is $2W=2M$. See
+\autoref{fig:suf2:extremal} for illustration on the extremal cases.
 
@@
 efficient on noise images than SUF without condensation, the mentioned
-consideration make the $2 \times 2$ optimization
-of SUF2 less dramatic in means of performance increase factor.
-Optimized SUF2 significantly outperforms optimized SUF only on
-huge noise images with less than 60\% density, while falling behind on
-small and medium-sized noise images.
+consideration make the $2 \times 2$ optimization of SUF2 less dramatic in means
+of performance increase factor. Optimized SUF2 significantly outperforms
+optimized SUF only on huge noise images with less than $60\%$ density, while
+falling behind on small and medium-sized noise images.
 
-Note that, while SUF2 is intended for CC counting, its memory
-consumption is only $O(W)$, which is significant for embedded systems and
-other resource-tight applications.
+Note that, while SUF2 is intended for CC counting, its memory consumption is
+only $O(W)$, which is significant for embedded systems and other resource-tight
+applications.
 
@@
 
-This modification of SUF was implemented to compare the optimization we present with
-another optimization suggested by Wu, Otoo and Suzuki\cite{wuotoo},
-namely decision tree.
-Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
-applicable for condensed image, as it exploits similar observations on 8-connectivity.
+This modification of SUF was implemented to compare the optimization we present
+with another optimization suggested by Wu, Otoo and Suzuki\cite{wuotoo}, namely
+decision tree. Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
+applicable for condensed image, as it exploits similar observations on
+8-connectivity.
 
-As before, the goal is to reduce the number of pixel lookup and label copy operations.
-This is achieved with careful consideration of only the necessary neighbors,
-stopping once the set of one or two label operations that are needed is
-determined. The decision tree for processing an object pixel is
-shown on \autoref{fig:dtsuf}.
+As before, the goal is to reduce the number of pixel lookup and label copy
+operations. This is achieved with careful consideration of only the necessary
+neighbors, stopping once the set of one or two label operations that are needed
+is determined. The decision tree for processing an object pixel is shown on
+\autoref{fig:dtsuf}.
 
@@
 
+\texten{
 For memory-tight applications such as embedded systems it can be useful to
 utilise the suggested condensation in a lazy way, so that without actually
-constructing the condensed image we use a function (``view'') calculating color of
-each cell with 4 lookups to the original binary image.
-This is probably most interesting when combined with SUF2 algorithm, as it is a way
-to win some time while avoiding memory footprint increase.
-However this lazy approach increases the number of image lookups.
+constructing the condensed image we use a function (``view'') calculating color
+of each cell with 4 lookups to the original binary image. This is probably most
+interesting when combined with SUF2 algorithm, as it is a way to win some time
+while avoiding memory footprint increase. However this lazy approach increases
+the number of image lookups.
 % So, according to our benchmark, this combination is efficient only for dense
 % ($>30\%$) noise.
+}\textru{
+В некоторых случаях может быть более приемлемо использовать сжатие неявно, то
+есть вместо построения 
+}
 
-So the benchmark charts show three levels of optimization applied to
-DFS, SUF and SUF2:
-just method name in legend refers to method applied to non-condensed image,
-``+view'' means lazy memory-less condensation
-and ``+2x2'' means that a temporary condensed image was constructed in memory.
-For the latter case, the construction time was included into the overall measured
-time of run.
+So the benchmark charts show three levels of optimization applied to DFS, SUF
+and SUF2: just method name in legend refers to method applied to non-condensed
+image, ``+view'' means lazy memory-less condensation and ``+2x2'' means that a
+temporary condensed image was constructed in memory. For the latter case, the
+construction time was included into the overall measured time of run.
 
@@
 
-The program was compiled using \texttt{g++ 5.3.0-3} compiler with \texttt{-O3} optimization flag.
-All tests were run on same machine running under Arch Linux with standard \texttt{linux 4.4.1-2} kernel,
-utilising non-overclocked \texttt{Intel(R) Core(TM) i5-2430M CPU @ 2.40GHz} (L3 cache size 3072K) processor and non-overclocked 1333MHz 8GiB Kingston RAM.
+The program was compiled using \texttt{g++ 5.3.0-3} compiler with \texttt{-O3}
+optimization flag. All tests were run on same machine running under Arch Linux
+with standard \texttt{linux 4.4.1-2} kernel, utilising non-overclocked
+\texttt{Intel(R) Core(TM) i5-2430M CPU @ 2.40GHz} (L3 cache size 3072K)
+processor and non-overclocked 1333MHz 8GiB Kingston RAM.
 
-The test images were generated in the following manner. At first, 
-for given density $d$, the first $4dNM$ pixels of
-the $2N \times 2M$ image were filled with object values, while rest filled with background values.
-Then all the pixels were permuted using \texttt{std::shuffle} function from \CXX standard library,
-which is an implementation of Fisher-Yates Shuffle algorithm\cite{fisher:yates}.
+The test images were generated in the following manner. At first, for given
+density $d$, the first $4dNM$ pixels of the $2N \times 2M$ image were filled
+with object values, while rest filled with background values. Then all the
+pixels were permuted using \texttt{std::shuffle} function from \CXX standard
+library, which is an implementation of Fisher-Yates Shuffle
+algorithm%\cite{fisher:yates}
+.
 
-All methods were run on same tests, which was achieved by supplying same seed to the
-\texttt{std::mt19937} Mersenne Twister\cite{mt19937} random number engine implementation from \CXX standard library before passing the engine to \texttt{std::shuffle}.
+All methods were run on same tests, which was achieved by supplying same seed
+to the \texttt{std::mt19937} Mersenne Twister\cite{mt19937} random number
+engine implementation from \CXX standard library before passing the engine to
+\texttt{std::shuffle}.
 
@@
 
-Line charts with benchmarks of mentioned algorithms on different noise images are shown on
+Line charts with benchmarks of mentioned algorithms on different noise images
+are shown on
 \Cref{fig:noise400x600,fig:noise800x1200,fig:noise1200x1800,fig:noise2000x3000,fig:noise4000x6000}.
 
-The extremely bigger CPU time for SUF and DTSUF methods on
-huge images could be explained by significant increase in the number of cache-misses
-produced by Union-Find.
-The charts prove effectiveness of the $2 \times 2$ condensation optimization.
+The extremely bigger CPU time for SUF and DTSUF methods on huge images could be
+explained by significant increase in the number of cache-misses produced by
+Union-Find. The charts prove effectiveness of the $2 \times 2$ condensation
+optimization.
 

### 2016-04-14 paper/paper.tex
Fix text width

@@
 characteristics referred to as linear variations.
-Kronrod%\cite{mak2}
+Kronrod\cite{mak2}
 introduced the notion of a regular component of the level set of a
@@
 library, which is an implementation of Fisher-Yates Shuffle
-algorithm%\cite{fisher:yates}
-.
+algorithm\cite{fisher:yates}.
 
@@
 \bibitem {mak1}
-Rudin, L., Osher, S., Fatemi, E.:
-Nonlinear total variation based noised removal algorithms.
-Phys. D. 60, 259-268 (1992)
+Rudin, L., Osher, S., Fatemi, E.: Nonlinear total variation based noised removal algorithms. Phys. D. 60, 259-268 (1992)
 
-% \bibitem {mak2}
-% Kronrod, A.:
-% On functions of two variables.
-% Uspehi Mat. Nauk 5, 1(35), 24-134(1950)
+\bibitem {mak2}
+Kronrod, A.: On functions of two variables ," Uspehi Mat. Nauk 5, 1(35), 24-134(1950)
 
 \bibitem {mak3}
-Makovetskii, A., Kober, V.:
-Image restoration based on topological properties of functions of two variables.
-Proc. SPIE Applications of Digital Image Processing XXXV, 8499, 84990A (2012)
+Makovetskii, A., Kober, V.: Image restoration based on topological properties of functions of two variables.  Proc. SPIE Applications of Digital Image Processing XXXV, 8499, 84990A (2012)
 
 \bibitem {mak4}
-Makovetskii, A., Kober, V.:
-Modified gradient descent method for image restoration.
-Proc. SPIE Applications of Digital Image Processing XXXVI, 8856, 885608-1 (2013)
+4.	Makovetskii, A., Kober, V.: Modified gradient descent method for image restoration. Proc. SPIE Applications of Digital Image Processing XXXVI, 8856, 885608-1 (2013)
 
 \bibitem {mak5}
-Chochia, P., Milukova, O.:
-Comparison of Two-Dimensional Variations in the Context of the Digital Image
-Complexity Assessment.
-Journal of Communications Technology and Electronics, 2015, 60, no. 12,  pp. 1432-1440
+Chochia, P., Milukova, O.: Comparison of Two-Dimensional Variations in the Context of the Digital Image Complexity Assessment.
+Journal of Communications Technology and Electronics, 2015, 60, no.
+12,  pp.
+1432-1440
 
@@
 He, L., Chao, Y., Suzuki, K., Wu, K.:
-Fast connected-component labeling.
-Pattern Recognition. v42, Pages 1977-1987. (2009)
+Fast connected-component labeling. Pattern Recognition. v42, Pages 1977-1987. (2009)
 
@@
 Cormen, H., Leiserson, E., Rivest, L., Stein, C.:
-Introduction to Algorithms (Second ed.).
-MIT Press (2001)
+Introduction to Algorithms (Second ed.), MIT Press (2001)
 
@@
 
-% \bibitem {fisher:yates}
-% Fisher, A., Yates, F.:
-% Statistical tables for biological, agricultural and medical research (3rd ed.).
-% London: Oliver \& Boyd. pp. 26–27. OCLC 14222135 (1948)
+\bibitem {fisher:yates}
+ Fisher, A., Yates, F.:
+ Statistical tables for biological, agricultural and medical research (3rd ed.). 
+ London: Oliver \& Boyd. pp. 26–27. OCLC 14222135 (1948)
 
@@
 Matsumoto, M., Nishimura, T.:
-Mersenne twister: a 623-dimensionally equidistributed uniform pseudo-random
-number generator.
+"Mersenne twister: a 623-dimensionally equidistributed uniform pseudo-random number generator".
 ACM Transactions on Modeling and Computer Simulation 8 (1): 3–30. (1998)

### 2016-04-19 paper/paper.tex

--- 2016-04-14 paper/paper.tex
@@
 %% раскомментировать следующие две команды
-% \renewcommand{\texten}[1]{}
-% \renewcommand{\textru}[1]{\foreignlanguage{russian}{\emph{#1}}{}}
+\renewcommand{\texten}[1]{}
+\renewcommand{\textru}[1]{\foreignlanguage{russian}{\emph{#1}}{}}
 
@@
 
-\texten{
 For memory-tight applications such as embedded systems it can be useful to
@@
 % ($>30\%$) noise.
-}\textru{
-В некоторых случаях может быть более приемлемо использовать сжатие неявно, то
-есть вместо построения 
-}
 
@@
 
-The test images were generated in the following manner. At first, for given
-density $d$, the first $4dNM$ pixels of the $2N \times 2M$ image were filled
-with object values, while rest filled with background values. Then all the
-pixels were permuted using \texttt{std::shuffle} function from \CXX standard
-library, which is an implementation of Fisher-Yates Shuffle
-algorithm%\cite{fisher:yates}
-.
+The test images were generated in the following manner. At first, 
+for given density $d$, the first $4dNM$ pixels of
+the $2N \times 2M$ image were filled with object values, while rest filled with background values.
+Then all the pixels were permuted using \texttt{std::shuffle} function from \CXX standard library,
+which is an implementation of Fisher-Yates Shuffle algorithm\cite{fisher:yates}.
 
@@
 
-Line charts with benchmarks of mentioned algorithms on different noise images
-are shown on
-\Cref{fig:overall:times}.
+Line charts with benchmarks of mentioned algorithms on different noise images are shown on
+\Cref{fig:noise400x600,fig:noise800x1200,fig:noise1200x1800,fig:noise2000x3000,fig:noise4000x6000}.
 
-The extremely bigger CPU time for SUF and DTSUF methods on huge images could be
-explained by significant increase in the number of cache-misses produced by
-Union-Find. The charts prove effectiveness of the $2 \times 2$ condensation
-optimization.
+The extremely bigger CPU time for SUF and DTSUF methods on
+huge images could be explained by significant increase in the number of cache-misses
+produced by Union-Find.
+The charts prove effectiveness of the $2 \times 2$ condensation optimization.
 
+\newcommand{\inputtable}[2]{
   \begin{figure}
     \centering
-  \includegraphics[height=0.45\textheight, keepaspectratio]{pics/overall.eps}
-  \caption{Average CPU time per run on random images $1200 \times 1800$ px.}
-  \label{fig:overall:times}
+    \includegraphics[height=0.45\textheight, keepaspectratio]{pics/#1x#2.eps}
+    \caption{Average CPU time per run on noise images $#1 \times #2$ px.}
+    \label{fig:noise#1x#2}
   \end{figure}
-
-\newcommand{\inputnorm}[1]{
-  \begin{subfigure}[t]{0.45\linewidth}
-    \centering
-    \includegraphics[width=\textwidth,
-                     keepaspectratio]{pics/normed_#1.eps}
-    \caption{Some.}
-    \label{fig:normed:#1}
-  \end{subfigure}
+ % \begin{table}[p]
+ %   \centering
+ %   \input{tables/#1x#2}
+ %   \caption{Average CPU time per run on noise images $#1 \times #2$ px}
+ %   \label{tab:noise#1x#2}
+ % \end{table}
 }
 
-\begin{figure}
-  \centering
-  \inputnorm{DFS}
-  \inputnorm{SUF}
-  \inputnorm{SUF2}
-  \inputnorm{DTSUF}
-  \label{fig:normed}
-\end{figure}
+% \inputtable{200}{300}
+\inputtable{400}{600}
+\inputtable{800}{1200}
+\inputtable{1200}{1800}
+\inputtable{2000}{3000}
+\inputtable{4000}{6000}
+% \inputtable{5000}{7500}
 

### 2016-04-19 paper/paper.tex
Add new normed pictures

@@
 %% раскомментировать следующие две команды
-\renewcommand{\texten}[1]{}
-\renewcommand{\textru}[1]{\foreignlanguage{russian}{\emph{#1}}{}}
+% \renewcommand{\texten}[1]{}
+% \renewcommand{\textru}[1]{\foreignlanguage{russian}{\emph{#1}}{}}
 

### 2016-04-19 utils/csv-to-latex.py

@@ -107,6 +107,7 @@ def plot_benchmarks(runs, out_file='output.eps', dpi=1200):
         lgd = plt.legend(loc=2, prop=fontP)
     else:
         lgd = plt.legend(loc=2, bbox_to_anchor=(1, 1), prop=fontP)
+    plt.grid(True)
     plt.savefig(out_file,
             bbox_extra_artists=(lgd,), bbox_inches='tight',
             format=out_file[-3:], dpi=dpi)

### 2016-04-19 utils/make_pics.py

@@ -1,0 +1,152 @@
+
+# coding: utf-8
+
+# In[1]:
+
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import pandas as pd
+
+import os
+
+
+# In[2]:
+
+pics_dir = os.path.join('paper', 'pics')
+
+
+# In[3]:
+
+data_path = os.path.join('bench-out', '1200x1800.csv')
+
+
+# In[4]:
+
+data = pd.read_csv(data_path)[['counter', 'grid', 'density', 'cpu time ms']].fillna('')
+
+
+# In[5]:
+
+COUNTER_MAP = {
+    "DecisionTreeDsuCounter": "DTSUF",
+    "DfsCounter": "DFS",
+    "DsuCounter": "SUF",
+    "TwolineDsuCounter": "SUF2"
+}
+data['counter'] = data['counter'].apply(lambda c: COUNTER_MAP.get(c) or c)
+
+
+# In[6]:
+
+GRID_MAP = {
+    'SimpleGrid': '',
+    'Viewing2x2Grid': ' on 2x2 view',
+    'Compressing2x2Grid': ' on 2x2 copy'
+}
+data['grid'] = data['grid'].apply(lambda c: GRID_MAP.get(c) or c)
+
+
+# In[7]:
+
+data['cg'] = data['counter'] + data['grid']
+
+
+# In[8]:
+
+piv = data.pivot(index='cg', columns='density', values='cpu time ms').swapaxes(0, 1)
+piv.head(2)
+
+
+# In[9]:
+
+cg = [(cou, gri) 
+      for cou in ('DFS', 'SUF', 'SUF2', 'DTSUF') 
+      for gri in ('', ' on 2x2 view', ' on 2x2 copy')
+      if (cou + gri in piv.columns)]
+
+
+# In[10]:
+
+METHOD_COLOR = {
+    'DFS': 'red',
+    'SUF': 'green',
+    "SUF2": 'blue',
+    'DTSUF': 'purple'
+}
+
+GRID_SHAPE = {
+    '': '-',
+    ' on 2x2 view': '--',
+    ' on 2x2 copy': '-.'
+}
+
+
+# In[67]:
+
+def add_plot(df, counter, grid):
+    name = counter + grid
+    if name in df.columns:
+        df[name].plot(style=GRID_SHAPE[grid], color=METHOD_COLOR[counter], legend=name)
+    
[…]

### 2016-04-19 utils/make_pics.py
Fix ylabel for normed pics

@@ -140,7 +140,7 @@ for cou in ('DFS', 'SUF', 'SUF2', 'DTSUF'):
     plot_all(
         norm_by(piv, cou)[[c + g for c, g in cg if c == cou or {c, cou} == {'SUF', 'DTSUF'}]],
         save='normed_' + cou,
-        ylabel='$T_A/T_{' + cou +  '}$',
+        ylabel='$T_{' + cou +  '}/T_A$',
         legend_loc=1 if cou == 'DTSUF' else 2,
         font_size='x-large')
     plt.clf()

### 2016-04-19 paper/paper.tex
Add caption to normed pics, clarify total time measurement

@@
 
-So the benchmark charts show three levels of optimization applied to
-DFS, SUF and SUF2:
-just method name in legend refers to method applied to non-condensed image,
-``+view'' means lazy memory-less condensation
-and ``+2x2'' means that a temporary condensed image was constructed in memory.
-For the latter case, the construction time was included into the overall measured
-time of run.
+So the benchmark charts show three levels of optimization applied to DFS, SUF
+and SUF2: just method name in legend refers to method applied to non-condensed
+image, `` on 2x2 view'' means lazy memory-less condensation and `` on 2x2 copy''
+means that a temporary condensed image was constructed in memory. For the latter
+case, the construction time was included into the overall measured time of run.
 
@@
 
-The program was compiled using \texttt{g++ 5.3.0-3} compiler with \texttt{-O3} optimization flag.
-All tests were run on same machine running under Arch Linux with standard \texttt{linux 4.4.1-2} kernel,
-utilising non-overclocked \texttt{Intel(R) Core(TM) i5-2430M CPU @ 2.40GHz} (L3 cache size 3072K) processor and non-overclocked 1333MHz 8GiB Kingston RAM.
-
-The test images were generated in the following manner. At first, 
-for given density $d$, the first $4dNM$ pixels of
-the $2N \times 2M$ image were filled with object values, while rest filled with background values.
-Then all the pixels were permuted using \texttt{std::shuffle} function from \CXX standard library,
-which is an implementation of Fisher-Yates Shuffle algorithm\cite{fisher:yates}.
-
-All methods were run on same tests, which was achieved by supplying same seed to the
-\texttt{std::mt19937} Mersenne Twister\cite{mt19937} random number engine implementation from \CXX standard library before passing the engine to \texttt{std::shuffle}.
+The program was compiled using \texttt{g++ 5.3.0-3} compiler with \texttt{-O3}
+optimization flag. All tests were run on same machine running under Arch Linux
+with standard \texttt{linux 4.4.1-2} kernel, utilising non-overclocked
+\texttt{Intel(R) Core(TM) i5-2430M CPU @ 2.40GHz} (L3 cache size 3072K)
+processor and non-overclocked 1333MHz 8GiB Kingston RAM.
+
+The test images were random images, generated in the following manner. At first,
+for given density $d$, the first $4dNM$ pixels of the $2N \times 2M$ image were
+filled with object values, while rest filled with background values. Then all
+the pixels were permuted using \texttt{std::shuffle} function from \CXX standard
+library, which is an implementation of Fisher-Yates Shuffle algorithm.
+
+All methods were run on same tests, which was achieved by supplying same seed to
+the \texttt{std::mt19937} Mersenne Twister\cite{mt19937} random number engine
+implementation from \CXX standard library before passing the engine to
+\texttt{std::shuffle}.
+
+It is worth noting again that for ``2x2 copy'' optimization, the apparent time
+overhead of constructing the extra condensed image  was included into the total
+run time, since we measured the total time between calling the CC counting
+subroutine, supplied with only the original image, and returning back from
+it.
 
@@
 Line charts with benchmarks of mentioned algorithms on different noise images
-are shown on
-\Cref{fig:overall:times}.
-
-The extremely bigger CPU time for SUF and DTSUF methods on huge images could be
-explained by significant increase in the number of cache-misses produced by
-Union-Find. The charts prove effectiveness of the $2 \times 2$ condensation
-optimization.
+are shown on \Cref{fig:overall}.
 
@@
   \caption{Average CPU time per run on random images $1200 \times 1800$ px.}
-  \label{fig:overall:times}
+  \label{fig:overall}
 \end{figure}
 
-\newcommand{\inputnorm}[1]{
+\newcommand{\inputnorm}[2]{
   \begin{subfigure}[t]{0.45\linewidth}
@@
                      keepaspectratio]{pics/normed_#1.eps}
-    \caption{Some.}
+    \caption{Speedup of #2}
     \label{fig:normed:#1}
@@
   \centering
-  \inputnorm{DFS}
-  \inputnorm{SUF}
-  \inputnorm{SUF2}
-  \inputnorm{DTSUF}
+  \inputnorm{DFS}{DFS optimizations}
+  \inputnorm{SUF}{SUF optimizations}
+  \inputnorm{SUF2}{SUF2 optimizations}
+  \inputnorm{DTSUF}{SUF optimizations, relative to DTSUF performance}
+  \caption{Speedup/density charts}
   \label{fig:normed}
@@
 \bibitem {mak1}
-Rudin, L., Osher, S., Fatemi, E.: Nonlinear total variation based noised removal algorithms. Phys. D. 60, 259-268 (1992)
-
-\bibitem {mak2}
-Kronrod, A.: On functions of two variables ," Uspehi Mat. Nauk 5, 1(35), 24-134(1950)
+Rudin, L., Osher, S., Fatemi, E.:
+Nonlinear total variation based noised removal algorithms.
+Phys. D. 60, 259-268 (1992)
+
+% \bibitem {mak2}
+% Kronrod, A.:
+% On functions of two variables.
+% Uspehi Mat. Nauk 5, 1(35), 24-134(1950)
 
 \bibitem {mak3}
-Makovetskii, A., Kober, V.: Image restoration based on topological properties of functions of two variables.  Proc. SPIE Applications of Digital Image Processing XXXV, 8499, 84990A (2012)
+Makovetskii, A., Kober, V.:
+Image restoration based on topological properties of functions of two variables.
+Proc. SPIE Applications of Digital Image Processing XXXV, 8499, 84990A (2012)
 
 \bibitem {mak4}
-4.	Makovetskii, A., Kober, V.: Modified gradient descent method for image restoration. Proc. SPIE Applications of Digital Image Processing XXXVI, 8856, 885608-1 (2013)
+Makovetskii, A., Kober, V.:
+Modified gradient descent method for image restoration.
+Proc. SPIE Applications of Digital Image Processing XXXVI, 8856, 885608-1 (2013)
 
 \bibitem {mak5}
-Chochia, P., Milukova, O.: Comparison of Two-Dimensional Variations in the Context of the Digital Image Complexity Assessment.
-Journal of Communications Technology and Electronics, 2015, 60, no.
-12,  pp.
-1432-1440
+Chochia, P., Milukova, O.:
+Comparison of Two-Dimensional Variations in the Context of the Digital Image
+Complexity Assessment.
+Journal of Communications Technology and Electronics, 2015, 60, no. 12,  pp. 1432-1440
 

### 2016-04-19 utils/make_pics.py

@@ -84,32 +84,37 @@ GRID_SHAPE = {
 
 # In[67]:
 
-def add_plot(df, counter, grid):
+def add_plot(df, counter, grid, legend_prefix=''):
     name = counter + grid
     if name in df.columns:
-        df[name].plot(style=GRID_SHAPE[grid], color=METHOD_COLOR[counter], legend=name)
-    
+        df[name].plot(
+                style=GRID_SHAPE[grid],
+                color=METHOD_COLOR[counter],
+                label=legend_prefix + name)
+
 def plot_all(df, font_size=None, ylabel=None, save=None, dpi=300, save_format='eps',
+            legend_prefix='',
             legend_on_right=False, legend_loc=None):
+
     from matplotlib.font_manager import FontProperties
 
     fontP = FontProperties()
     if font_size:
         fontP.set_size(font_size)
-    
+
     for counter, grid in cg:
-        add_plot(df, counter=counter, grid=grid)
-        
+        add_plot(df, counter=counter, grid=grid, legend_prefix=legend_prefix)
+
     plt.grid(True)
     if legend_on_right:
         lgd = plt.legend(loc=2, bbox_to_anchor=(1.01, 1), prop=fontP)
     else:
         lgd = plt.legend(loc=legend_loc, prop=fontP)
-        
+
     plt.xlabel('density, %')
     if ylabel:
         plt.ylabel(ylabel)
-    
+
     if save:
         plt.savefig(os.path.join(pics_dir, save + '.' + save_format),
             bbox_extra_artists=(lgd,), bbox_inches='tight',
@@ -141,6 +146,7 @@ for cou in ('DFS', 'SUF', 'SUF2', 'DTSUF'):
         norm_by(piv, cou)[[c + g for c, g in cg if c == cou or {c, cou} == {'SUF', 'DTSUF'}]],
         save='normed_' + cou,
         ylabel='$T_{' + cou +  '}/T_A$',
+        legend_prefix='A = ',
         legend_loc=1 if cou == 'DTSUF' else 2,
         font_size='x-large')
     plt.clf()

### 2016-04-19 paper/paper.tex
Add results explanation

@@
 namely decision tree.
-Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
-applicable for condensed image, as it exploits similar observations on 8-connectivity.
+Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not applicable for
+condensed image, as it exploits similar observations on 8-connectivity.
 }\textru{
@@
 
+Benchmarks showed that for random images, the run time depends non-trivially on
+the value of density $d$, which is the ratio of object pixels in the image. On
+the one hand, this is because all of the considered algorithms pass background
+pixels without any processing, and somehow process the object pixels, so the
+exact number of operations such as Union and Find should generally grow with
+$d$. On the other hand, the greater the value of $d$, the less number of CC is
+likely to be present in the image, which reduces the number of relatively
+expensive Union operations.
+
+So benchmarks were run for several values of $d$ to achieve a picture of how the
+speedup depends on the number of object pixels in the image.
+
 It is worth noting again that for ``2x2 copy'' optimization, the apparent time
@@
 
-Line charts with benchmarks of mentioned algorithms on different noise images
-are shown on \Cref{fig:overall}.
+The overall benchmark line plot of all algorithms and optimizations is
+presented on \Cref{fig:overall}.
 
@@
 
+To show more closely how efficient the optimizations actually appear to be, we
+plotted speedup/density charts for the considered methods (\Cref{fig:normed}).
+Here the speedup of algorithm A relative to algorithm B is defined as
+$\frac{T_B}{T_A}$, where $T_A$ and $T_B$ are the total times used by the
+corresponding algorithms to count CC in all the test images of same density.
+The greater the speedup of optimized method relative to original method, the
+more efficient the optimization is considered.
+
 \newcommand{\inputnorm}[2]{
@@
   \inputnorm{SUF2}{SUF2 optimizations}
-  \inputnorm{DTSUF}{SUF optimizations, relative to DTSUF performance}
+  \inputnorm{DTSUF}{SUF optimizations, relative to DTSUF}
   \caption{Speedup/density charts}
@@
 \end{figure}
+
+The speedup plots on \Cref{fig:normed:DFS,fig:normed:SUF,fig:normed:SUF2}
+suggest that the ``2x2 copy'' optimization speeds up the considered methods by
+factor greater than 1 for all tested densities. On the other hand, the
+``2x2 view'' optimization is showed to be not time-efficient at all for some
+densities, slowing down DFS and SUF2 if run on images with densities less than
+38\% and 28\% respectively. Utilizing the latter optimization also does not
+speed up SUF on images on density approaching 20\%, although it does not slow
+down the algorithms either.
+
+The worst-case and best-case speedups are shown on the \autoref{tab:speedup}.
+Here we see that while ``2x2 view'' optimization is more memory-efficient than
+``2x2 copy'', it can slow down the method in some situations, and generally it
+does not give as much performance advantage.
+
+We should also mention the chart on \autoref{fig:normed:DTSUF}, featuring
+speedup of our SUF optimizations relative to DTSUF in order to compare our
+optimizations to the optimization by Wu, Otoo and Suzuki. The plot clearly shows
+that for all the densities under 80\% our ``SUF on 2x2 copy'' algorithm
+outperforms the decision-tree based approach by the factor of at least
+$1.5$, with speedup reaching $2.5$ for 2\% density.
+
+\begin{table}
+  \centering
+  \begin{tabular}{l|c|c|c|c}
+    Algorithm & worst speedup & worst speedup density & best speedup & best
+    speedup density \\
+    \hline
+    \input{tables/speedup}
+  \end{tabular}
+  \caption{The worst and the worst speedups of the considered optimizations}
+  \label{tab:speedup}
+\end{table}
 

### 2016-04-19 utils/make_pics.py

@@ -61,7 +61,7 @@ piv.head(2)
 # In[9]:
 
 cg = [(cou, gri) 
-      for cou in ('DFS', 'SUF', 'SUF2', 'DTSUF') 
+      for cou in ('DFS', 'SUF', 'DTSUF', 'SUF2')
       for gri in ('', ' on 2x2 view', ' on 2x2 copy')
       if (cou + gri in piv.columns)]
 

### 2016-04-20 paper/paper.tex
Merge Mikhail's changes

@@
 \usepackage{graphicx}
+
+\usepackage{verbatim}
 
@@
 \begin{abstract}
-The linear variation is a topological characteristic of the function of two variables.
-The problem of linear variation computing can be reduced to the problem of counting connected components in a binary image with eight-connected connectivity.
+The linear variation is a topological characteristic of the function of two
+variables. The problem of linear variation computing can be reduced to the
+problem of counting connected components in a binary image with eight-connected
+connectivity.
 %The new hybrid method for the problem is presented.
@@
 %that groups the pixels into $2 \times 2$ cells.
-This is usually done through connected components labeling, and many approaches for that are known.
-This paper presents an efficient way to convert the initial binary image with 8-connectivity into a condensed (non-binary) image with new connectivity that is 2 times smaller in each dimension than the initial image.
-Most approaches for connected components labeling are still valid for the new image.
-As time and memory consumptions of conventional approaches usually depend linearly on the number of pixels,
-running them on the condensed image can be up to 4 times more efficient than running on the initial image.
-The method is essentially a modification of some known raster algorithms for labeling that groups pixels into $2 \times 2$ cells.
-A performance benchmark of the proposed  method applied to some approaches on noise images is provided.
-
-\keywords{Binary raster image, connected component labeling, pattern recognition}
+%This is usually done through connected components labeling, and many approaches
+%for that are known. This paper presents an efficient way to convert the initial
+%binary image with 8-connectivity into a condensed %(non-binary) image with new
+%connectivity that is 2 times smaller in each dimension than the initial image.
+%Most approaches for connected components labeling are still valid for the new
+%image. As time and memory consumptions of conventional approaches usually
+%depend linearly on the number of pixels, running them on the condensed image
+%can be up to 4 times more efficient than running on the initial image.
+The method is essentially a modification of some known raster algorithms for
+labeling that groups pixels into $2 \times 2$ cells. A performance benchmark of
+the proposed  method applied to some approaches on noise images is provided.
+
+\keywords{Binary raster image, connected component labeling, pattern
+recognition}
 \end{abstract}
@@
 
+\begin{comment}
 In image restoration it is often necessary to consider the following problem.
@@
 \end{equation}
-where $\norm{\cdot}$ is the $L_2$ norm and all function we consider belong to the class
-$BV(\Omega)$, i.\,e. the class of bounded variation on the set $\Omega$ functions.
+where $\norm{\cdot}$ is the $L_2$ norm and all function we consider belong to
+the class $BV(\Omega)$, i.\,e. the class of bounded variation on the set
+$\Omega$ functions.
 %
@@
 
-The norm and modulus of the gradient are metrical characteristic of a function of
-two variables.
+The norm and modulus of the gradient are metrical characteristic of a function
+of two variables.
 Continuous functions of two variables also have a set of topological
@@
 continuous function of two variables.
-The simplest topological characteristic in the linear variation theory is a number
-of regular components for all level sets of a function.
+The simplest topological characteristic in the linear variation theory is a
+number of regular components for all level sets of a function.
 Full information about linear variations of a function is contained in the
@@
 
-Let $w$ be a binary discrete function $w=(w_{i,j})$, where $w_{i,j} \in \{0,1\}$,
-for all pairs $(i,j)$.
-A subset of such pairs $(i, j)$ when $w_{i,j}=1$ and all elements of the subset are
-connected by the 8-connectivity, is called the connected component of the binary
-function $w$.
+Let $w$ be a binary discrete function $w=(w_{i,j})$, where $w_{i,j} \in
+\{0,1\}$, for all pairs $(i,j)$. A subset of such pairs $(i, j)$ when
+$w_{i,j}=1$ and all elements of the subset are connected by the 8-connectivity,
+is called the connected component of the binary function $w$.
 For a number $k \in \N$ and a discrete function $u$ we define the following
@@
 \begin{definition}
-  The linear variation $V(u_{i,j})$ of a discrete function  is defined as follows:
+  The linear variation $V(u_{i,j})$ of a discrete function  is defined as
+  follows:
   \begin{equation}
@@
 
-Suppose that if the pair $(i,j)$ is outside of the domain of the function $u$, then 
-$u_{i,j} = 0$.
+Suppose that if the pair $(i,j)$ is outside of the domain of the function $u$,
+then $u_{i,j} = 0$.
 The main problem of the computer implementation of the linear variation is 
@@
 
+\end{comment}
+
+Continuous functions of two variables have a set of topological characteristics
+referred to as linear variations. Kronrod\cite{mak2} introduced the notion of a
+regular component of the level set of a continuous function of two variables.
+The simplest topological characteristic in linear variation theory is the number
+of regular components for all level sets of a function. In \cite{mak3,mak4,mak5}
+was discussed the necessity of the using a linear variation in the image
+restoration theory.
+
+The problem of computing the linear variation in discrete case is equivalent 
+to the problem of counting the number of connected components (CC) in a binary
+image with 8-connectivity. This is usually done through CC labeling, and many
+approaches for that are known\cite{hechao}. We came to the labeling problem
+while searching the algorithm for linear variation computing. Of course, the
+labeling problem also is well known in the digital image processing theory.
+
 % Alekseev part
 
-This problem is equivalent to the problem of counting the number of connected
-components (CC) in a binary image with 8-connectivity.
-This is usually done through CC labeling, and many approaches for that are
-known\cite{hechao}.
-
-This paper presents an efficient way to convert the initial binary image
-with 8-connectivity into
-a condensed (non-binary) image with new connectivity that is $2$ times smaller in
-each dimension than the initial image.
-Most approaches for CC labeling are still valid for the new image.
-As time and memory consumptions of conventional approaches usually depend linearly on
-the number of pixels, running them on the condensed image can be
-more efficient than running on the initial image.
-
-It is worth noting that our method exploits an important property of 8-connectivity
-that does not hold for 4-connectivity, so it is not applicable for the latter case.
+This paper presents an efficient way to convert the initial binary image with
+8-connectivity into a condensed (non-binary) image with new connectivity that is
+$2$ times smaller in each dimension than the initial image. Most approaches for
+CC labeling are still valid for the new image. As time and memory consumptions
+of conventional approaches usually depend linearly on the number of pixels,
+running them on the condensed image can be more efficient than running on the
+initial image.
+
+It is worth noting that our method exploits an important property of
+8-connectivity that does not hold for 4-connectivity, so it is not applicable
+for the latter case.
 
@@
 We will use the fact that for the case of 8-connectivity all $4$ pixels of any
-$2 \times 2$ square are adjacent to each other, and so are contained in the same CC
-and should have same labels upon completion of the labeling algorithm.
-So instead of labeling single pixels we can label $2 \times 2$ groups of pixels,
-or ``cells''.
+$2 \times 2$ square are adjacent to each other, and so are contained in the same
+CC and should have same labels upon completion of the labeling algorithm. So
+instead of labeling single pixels we can label $2 \times 2$ groups of pixels, or
+``cells''.
 }\textru{
 Наблюдение: с 8-связностью в любом квадратике $2 \times 2$ любые два пикселя
-находятся рядом. Значит если это пара объектных пикселей, то они находятся в одной
-КС, и значит каждый такой квадратик пересекается не более чем с одной КС.
+находятся рядом. Значит если это пара объектных пикселей, то они находятся в
+одной КС, и значит каждый такой квадратик пересекается не более чем с одной КС.
 Поэтому вместо маркировки отдельных пикселей можно маркировать такие квадраты.
@@
 }
-specified by a binary predicate $b(x, y)$,
-returning $1$ in case the pixel at the given coordinates is black (object pixel),
-and $0$ otherwise (background pixel).
-For convenience we use 0-indexation throughout the paper, so the topmost row has
-row index $y=0$, and the leftmost column has column index $x=0$.
+specified by a binary predicate $b(x, y)$, returning $1$ in case the pixel at
+the given coordinates is black (object pixel), and $0$ otherwise (background
+pixel). For convenience we use 0-indexation throughout the paper, so the topmost
+row has row index $y=0$, and the leftmost column has column index $x=0$.
 }\textru{
@@
 \texten{
-The condensed image will be of size $N \times M$.
-Each pixel of the condensed image will unambiguously represent the configuration
-of the corresponding $2\times 2$ cell of pixels of the initial binary image.
-So the condensed image will contain pixels of $2^4 = 16$ colors.
-We suggest the function $c(x, y)$ denoting the color of a pixel of condensed image
-to be the following:
-}\textru{
-Сжатое изображение будет иметь размер $N \times M$.
-Каждый пиксель полученного изображения должен будет однозначно задавать конфигурацию
-соответствующих четырёх пикселей исходного изображения, поэтому он может быть одного
-из $2^4 = 16$ цветов. Предлагается такая функция $c(x,y)$, кодирующая четыре исходных
+The condensed image will be of size $N \times M$. Each pixel of the condensed
+image will unambiguously represent the configuration of the corresponding
+$2\times 2$ cell of pixels of the initial binary image. So the condensed image
+will contain pixels of $2^4 = 16$ colors. We suggest the function $c(x, y)$
+denoting the color of a pixel of condensed image to be the following:
+}\textru{
+Сжатое изображение будет иметь размер $N \times M$. Каждый пиксель полученного
+изображения должен будет однозначно задавать конфигурацию соответствующих
+четырёх пикселей исходного изображения, поэтому он может быть одного из $2^4 =
+16$ цветов. Предлагается такая функция $c(x,y)$, кодирующая четыре исходных
 пикселя в цвет клетки:
@@
 So the value of each pixel of the initial image is stored in the corresponding
-bit in color of the corresponding pixel of the condensed image.
-Note that if $c(x, y) = 0$ for some $(x, y)$, then there are no object pixels in
-this cell, and no label should be assigned to this whole cell. The enumeration of
-pixels inside one cell is shown on \autoref{fig:pixelsEnumeration}.
-}\textru{
-Таким образом значение каждого из исходных пикселей просто записано в соответствующий
-бит в итоговом пикселе. Важно отметить, что если все четыре исходных пикселя были
-фоновые, то и клетка будет иметь цвет 0. Нумерация пикселей внутри клетки показана на
-\autoref{fig:pixelsEnumeration}.
-}
-
-\begin{figure}[t]
+bit in color of the corresponding pixel of the condensed image. Note that if
+$c(x, y) = 0$ for some $(x, y)$, then there are no object pixels in this cell,
+and no label should be assigned to this whole cell. The enumeration of pixels
+inside one cell is shown on \autoref{fig:pixelsEnumeration}.
+}\textru{
+Таким образом значение каждого из исходных пикселей просто записано в
+соответствующий бит в итоговом пикселе. Важно отметить, что если все четыре
+исходных пикселя были фоновые, то и клетка будет иметь цвет 0. Нумерация
+пикселей внутри клетки показана на \autoref{fig:pixelsEnumeration}.
+}
+
+\begin{figure}
   \centering
@@
 \texten{
-However, for every variant of relative placement of two nearby cells there is
-a mask of pixels of each cells that are important for direct connectivity of
-this pair of cells.
-So two cells are considered directly connected, if they are nearby
-in terms of 8-connectivity and each of them contains at least one object pixel
-in the mask of important pixels corresponding to relative placement of the cells.
-The masks are shown in \autoref{fig:connectivity:masks}.
-}\textru{
-Тем не менее, можно заметить, что анализировать можно не все биты цвета, а только
-некоторую маску тех бит, которые важны для связности двух конкретных клеток.
-Эти маски показаны на \autoref{fig:connectivity:masks}.
+However, for every variant of relative placement of two nearby cells there is a
+mask of pixels of each cells that are important for direct connectivity of this
+pair of cells. So two cells are considered directly connected, if they are
+nearby in terms of 8-connectivity and each of them contains at least one object
+pixel in the mask of important pixels corresponding to relative placement of the
+cells. The masks are shown in \autoref{fig:connectivity:masks}.
+}\textru{
+Тем не менее, можно заметить, что анализировать можно не все биты цвета, а
+только некоторую маску тех бит, которые важны для связности двух конкретных
+клеток. Эти маски показаны на \autoref{fig:connectivity:masks}.
 }
@@
   \caption{Masks of important pixels for all four cases of relative placement
-  of two nearby cells. Those pixels are filled in gray, with their numbers within
-  cell specified.}
+    of two nearby cells. Those pixels are filled in gray, with their numbers
+    within cell specified.}
   \label{fig:connectivity:masks}
@@
 operations.
-As an example, \autoref{fig:connectivity:code} compares possible codes in \C checking
-if the labels of the current pixel or cell and the one right on top of it
-should be same for classic 8-connectivity and for the new connectivity.
-Note that as bitwise operations are usually relatively cheap, we added only a
-little complication compared to decreasing the number of pixels to process by the 
+As an example, \autoref{fig:connectivity:code} compares possible codes in \C
+checking if the labels of the current pixel or cell and the one right on top of
+it should be same for classic 8-connectivity and for the new connectivity. Note
+that as bitwise operations are usually relatively cheap, we added only a little
+complication compared to decreasing the number of pixels to process by the
 factor of 4.
@@
 ли текущий пиксель (или клетка) с пикселем (или клеткой) над ним (ней).
-Стоит отметить, что хотя конструкция
-усложнилась, битовые операции как правило очень эффективны, и это усложнение, скорее
-всего, незначительно по сравнению с уменьшением количества рассматриваемых позиций в 4
-раза.
+Стоит отметить, что хотя конструкция усложнилась, битовые операции как правило
+очень эффективны, и это усложнение, скорее всего, незначительно по сравнению с
+уменьшением количества рассматриваемых позиций в 4 раза.
 }
@@
 }\textru{
-Так как все рассматриваемые методы, кроме DTSUF, можно применять как для сжатого,
-так и для несжатого изображения, мы обозначим размеры входного изображения
-каждого изображения за $H$ и $W$.
+Так как все рассматриваемые методы, кроме DTSUF, можно применять как для
+сжатого, так и для несжатого изображения, мы обозначим размеры входного
+изображения каждого изображения за $H$ и $W$.
 }
@@
 Also to avoid repeating ``pixel or cell'' we will just use the word ``pixel''
-meaning the appropriate raster for non-condensed and condensed images.
-When we mention neighbors of some pixel, we mean pixels that are directly connected
-to it in terms of appropriate connectivity.
+meaning the appropriate raster for non-condensed and condensed images. When we
+mention neighbors of some pixel, we mean pixels that are directly connected to
+it in terms of appropriate connectivity.
 }\textru{
@@
 
-The first algorithm we chose for the benchmark is Depth First Search graph traversal.
-Graph traversal algorithms are the general approach for CC labeling in graphs, and
+The first algorithm we chose for the benchmark is Depth First Search graph
+traversal. Graph traversal algorithms are the general approach for CC labeling
+in graphs, and
 DFS is arguably the most simple.
@@
 
-Like most approaches, the algorithm has an outer loop over all pixels of the image.
-Each time an object pixel (non-empty cell, in case it is actually a cell)
-which is not labeled yet is encountered, DFS traversal is launched from that pixel.
-The traversal labels all the object pixels of the same CC, launching recursively
-from all unlabeled neighbors of current pixel.
+Like most approaches, the algorithm has an outer loop over all pixels of the
+image. Each time an object pixel (non-empty cell, in case it is actually a cell)
+which is not labeled yet is encountered, DFS traversal is launched from that
+pixel. The traversal labels all the object pixels of the same CC, launching
+recursively from all unlabeled neighbors of current pixel.
 
@@
 Another general approach for graph CC labeling which is more popular in image
-labeling is using some kind of a data structure to maintain disjoint sets of pixels
-that are considered to be in the same CC.
+labeling is using some kind of a data structure to maintain disjoint sets of
+pixels that are considered to be in the same CC.
 }\textru{
@@
 \texten{
-The structure is commonly called Union-Find, as the two essential operations
-it is supposed to support are Union, which is used to merge some two of the
+The structure is commonly called Union-Find, as the two essential operations it
+is supposed to support are Union, which is used to merge some two of the
 disjoint sets in one, and Find, which is used to find the representative element
-of the set some element belongs to, so that we can always tell if two elements are
-in the same set.
-The structure can be implemented so that the amortized time complexity for each
-operation is $O(\alpha(HW))$, where $\alpha$ is the inverse Ackermann function,
-and $\alpha(HW) < 5$ for all remotely practical image sizes\cite{CLRS}. % TODO links
+of the set some element belongs to, so that we can always tell if two elements
+are in the same set. The structure can be implemented so that the amortized time
+complexity for each operation is $O(\alpha(HW))$, where $\alpha$ is the inverse
+Ackermann function, and $\alpha(HW) < 5$ for all remotely practical image
+sizes\cite{CLRS}. % TODO links
 }\textru{
@@
 $\alpha$~--- обратная функция Аккермана.
-Для всех практически возможных размеров изображения эта функция не превосходит 5.
+Для всех практически возможных размеров изображения эта функция не превосходит
+5.
 }
@@
 \texten{
-At first each object pixel is considered to belong to its own CC, so right before
-considering the neighbors we increment the CC counter.
-Moreover, it is associated with its own disjoint set in the Union-Find structure.
-More specifically, object pixel at position $x, y$ is associated with
-the set number $xW + y$ to guarantee absence of collisions.
+At first each object pixel is considered to belong to its own CC, so right
+before considering the neighbors we increment the CC counter. Moreover, it is
+associated with its own disjoint set in the Union-Find structure. More
+specifically, object pixel at position $x, y$ is associated with the set number
+$xW + y$ to guarantee absence of collisions.
 }\textru{
@@
 независимым множеством в структуре Union-Find, состоящем из одного элемента.
-Конкретно пиксель в позиции $x, y$ заранее ассоциирован с элементом номер $xW + y$,
-чтобы гарантировать отсутствие коллизий.
-}
-
-\texten{
-The possible neighbors are processed in the order shown on \autoref{fig:suf:neighbors},
-and for each neighbor that is object pixel, the Union operation is performed
-on the two sets containing the current and the neighbor pixel.
-In case the Union operation was successful, which means that the pixels
-had belonged previously to different sets and we have just found a way to merge two
-CC, the CC counter is decremented.
+Конкретно пиксель в позиции $x, y$ заранее ассоциирован с элементом номер $xW +
+y$, чтобы гарантировать отсутствие коллизий.
+}
+
+\texten{
+The possible neighbors are processed in the order shown on
+\autoref{fig:suf:neighbors}, and for each neighbor that is object pixel, the
+Union operation is performed on the two sets containing the current and the
+neighbor pixel. In case the Union operation was successful, which means that the
+pixels had belonged previously to different sets and we have just found a way to
+merge two CC, the CC counter is decremented.
 }\textru{
@@
 \texten{
-Note that this algorithm outputs not the matrix of final labels, but only the number
-of CC.
-This is the place in which it differs from some common two-pass approaches for CC % TODO references
-labeling, as it is an algorithm not for labeling, but just for counting CC.
-The latter is exactly what is needed for the aforementioned linear variation problem.
+Note that this algorithm outputs not the matrix of final labels, but only the
+number of CC. This is the place in which it differs from some common two-pass
+approaches for CC labeling, as it is an algorithm not for labeling, but just for
+counting CC. The latter is exactly what is needed for the aforementioned linear
+variation problem.
 }\textru{
@@
 \texten{
-Also we have to mention that for performance reasons the Union-Find internal structures
-are allocated ahead to have the size of at least $HW$ to guarantee that
-each object can be easily associated with unique starting set given only its
-coordinates. So memory consumption of SUF is linear in the number of pixels.
+Also we have to mention that for performance reasons the Union-Find internal
+structures are allocated ahead to have the size of at least $HW$ to guarantee
+that each object can be easily associated with unique starting set given only
+its coordinates. So memory consumption of SUF is linear in the number of pixels.
 }\textru{
@@
 \texten{
-The idea is to run through each row twice: first time to establish label equivalence
-classes within row, and the second time to set unique equal label for all pixels of
-same CC.
-This gives us the opportunity to maintain just $O(W)$
-different labels at each moment, and so the Union-Find structure can be allocated
-for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.
+The idea is to run through each row twice: first time to establish label
+equivalence classes within row, and the second time to set unique equal label
+for all pixels of same CC. This gives us the opportunity to maintain just $O(W)$
+different labels at each moment, and so the Union-Find structure can be
+allocated for only $O(W)$ elements, as opposed to $WH$ elements needed for SUF.
 }\textru{
@@
 The exact maximum number of label equivalence classes between two adjacent rows
-depends on whether the given image is condensed.
-For non-condensed images $W=2M$, and maximum number of label equivalence classes
-in two adjacent rows is $M=\frac{W}2$, as if two object pixels are both in adjacent
-rows and in adjacent columns, then they are neighbors in terms of 8-connectivity.
+depends on whether the given image is condensed. For non-condensed images
+$W=2M$, and maximum number of label equivalence classes in two adjacent rows is
+$M=\frac{W}2$, as if two object pixels are both in adjacent rows and in adjacent
+columns, then they are neighbors in terms of 8-connectivity.
 }\textru{
@@
 \texten{
-On the other hand, for condensed images $W=M$, and maximum number of label equivalence
-class in two adjacent rows of cells is $2W=2M$.
-See \autoref{fig:suf2:extremal} for illustration on the extremal cases.
+On the other hand, for condensed images $W=M$, and maximum number of label
+equivalence class in two adjacent rows of cells is $2W=2M$. See
+\autoref{fig:suf2:extremal} for illustration on the extremal cases.
 }\textru{
@@
 \texten{
-This modification of SUF was implemented to compare the optimization we present with
-another optimization suggested by Wu, Otoo and Suzuki\cite{wuotoo},
-namely decision tree.
-Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not applicable for
-condensed image, as it exploits similar observations on 8-connectivity.
+This modification of SUF was implemented to compare the optimization we present
+with another optimization suggested by Wu, Otoo and Suzuki\cite{wuotoo}, namely
+decision tree. Unfortunately DTSUF (Decision Tree Scan plus Union-Find) is not
+applicable for condensed image, as it exploits similar observations on
+8-connectivity.
 }\textru{
@@
 \texten{
-As before, the goal is to reduce the number of pixel lookup and label copy operations.
-This is achieved with careful consideration of only the necessary neighbors,
-stopping once the set of one or two label operations that are needed is
-determined. The decision tree for processing an object pixel is
-shown on \autoref{fig:dtsuf}.
+As before, the goal is to reduce the number of pixel lookup and label copy
+operations. This is achieved with careful consideration of only the necessary
+neighbors, stopping once the set of one or two label operations that are needed
+is determined. The decision tree for processing an object pixel is shown on
+\autoref{fig:dtsuf}.
 }\textru{
@@
 utilise the suggested condensation in a lazy way, so that without actually
-constructing the condensed image we use a function (``view'') calculating color of
-each cell with 4 lookups to the original binary image.
-This is probably most interesting when combined with SUF2 algorithm, as it is a way
-to win some time while avoiding memory footprint increase.
-However this lazy approach increases the number of image lookups.
+constructing the condensed image we use a function (``view'') calculating color
+of each cell with 4 lookups to the original binary image. This is probably most
+interesting when combined with SUF2 algorithm, as it is a way to win some time
+while avoiding memory footprint increase. However this lazy approach increases
+the number of image lookups.
 % So, according to our benchmark, this combination is efficient only for dense
@@
 
-For measurement purposes all algorithms and optimizations were implemented by us in \CXX.
+For measurement purposes all algorithms and optimizations were implemented by us
+in \CXX.
 
@@
 that for all the densities under 80\% our ``SUF on 2x2 copy'' algorithm
-outperforms the decision-tree based approach by the factor of at least
-$1.5$, with speedup reaching $2.5$ for 2\% density.
+outperforms the decision-tree based approach by the factor of at least $1.5$,
+with speedup reaching $2.5$ for 2\% density.
 
@@
 
-% \section{Summary and future work}
+\section{Conclusion}
+
+The problem of linear variation computing can be reduced to the problem of
+counting connected components in a binary image with eight-connected
+connectivity. This paper presents an efficient way to convert the initial binary
+image with 8-connectivity into a condensed (non-binary) image with new
+connectivity that is 2 times smaller in each dimension than the initial image.
+Most approaches for connected components labeling are still valid for the new
+image. As time and memory consumptions of conventional approaches usually depend
+linearly on the number of pixels, running them on the condensed image can be up
+to 4 times more efficient than running on the initial image. The method is
+essentially a modification of some known raster algorithms for labeling that
+groups pixels into $2 \times 2$ cells. A performance benchmark of the proposed
+method applied to some approaches on noise images is provided. It is shown at
+least up to 2 times more quick results could be received.
+
+\section{Acknowledgements}
+
+The work was supported by the Ministry of Education and Science of Russian
+Federation (grant 2/1766/2014K).
 
@@
 
-\bibitem {mak1}
-Rudin, L., Osher, S., Fatemi, E.:
-Nonlinear total variation based noised removal algorithms.
-Phys. D. 60, 259-268 (1992)
-
-% \bibitem {mak2}
-% Kronrod, A.:
-% On functions of two variables.
-% Uspehi Mat. Nauk 5, 1(35), 24-134(1950)
+%\bibitem {mak1}
+% Rudin, L., Osher, S., Fatemi, E.:
+% Nonlinear total variation based noised removal algorithms.
+% Phys. D. 60, 259-268 (1992)
+
+\bibitem {mak2}
+Kronrod, A.:
+On functions of two variables.
+Uspehi Mat. Nauk 5, 1(35), 24-134(1950)
 
@@
 Complexity Assessment.
-Journal of Communications Technology and Electronics, 2015, 60, no. 12,  pp. 1432-1440
+Journal of Communications Technology and Electronics, 2015, 60, no. 12,  pp.
+1432-1440
 

### 2016-04-20 paper/paper.tex
Some proofreading, make conclusion more concrete.

@@
 \usepackage{graphicx}
+
+\usepackage{braket}
 
@@
 \begin{abstract}
-The linear variation is a topological characteristic of the function of two
+Linear variation is a topological characteristic of a function of two
 variables. The problem of linear variation computing can be reduced to the
@@
 connectivity.
-%The new hybrid method for the problem is presented.
-%The method is essentially a modification of a two-scan algorithm for labeling
-%that groups the pixels into $2 \times 2$ cells.
-%This is usually done through connected components labeling, and many approaches
-%for that are known. This paper presents an efficient way to convert the initial
-%binary image with 8-connectivity into a condensed %(non-binary) image with new
-%connectivity that is 2 times smaller in each dimension than the initial image.
-%Most approaches for connected components labeling are still valid for the new
-%image. As time and memory consumptions of conventional approaches usually
-%depend linearly on the number of pixels, running them on the condensed image
-%can be up to 4 times more efficient than running on the initial image.
-The method is essentially a modification of some known raster algorithms for
-labeling that groups pixels into $2 \times 2$ cells. A performance benchmark of
-the proposed  method applied to some approaches on noise images is provided.
+The proposed method is essentially a modification of some known raster
+algorithms for connected components labeling that groups pixels into $2 \times
+2$ cells. A performance benchmark of the proposed method applied to some
+approaches on random images is provided.
 
@@
 Continuous functions of two variables have a set of topological characteristics
-referred to as linear variations. Kronrod\cite{mak2} introduced the notion of a
-regular component of the level set of a continuous function of two variables.
+referred to as linear variations. Kronrod\cite{mak2} introduced the notion of
+regular component of the level set of continuous function of two variables.
 The simplest topological characteristic in linear variation theory is the number
-of regular components for all level sets of a function. In \cite{mak3,mak4,mak5}
-was discussed the necessity of the using a linear variation in the image
+of regular components for all level sets of a function. Works
+\cite{mak3,mak4,mak5} address the necessity of linear variation in image
 restoration theory.
 
-The problem of computing the linear variation in discrete case is equivalent 
-to the problem of counting the number of connected components (CC) in a binary
-image with 8-connectivity. This is usually done through CC labeling, and many
+The problem of computing linear variation in discrete case is equivalent to the
+problem of counting the number of connected components (CC) in a binary image
+with 8-connectivity. This is usually done through CC labeling, and many
 approaches for that are known\cite{hechao}. We came to the labeling problem
-while searching the algorithm for linear variation computing. Of course, the
-labeling problem also is well known in the digital image processing theory.
+while searching an efficient algorithm algorithm for linear variation
+computation. Of course, the labeling problem also is well known in the digital
+image processing theory.
 
@@
 It is worth noting that our method exploits an important property of
-8-connectivity that does not hold for 4-connectivity, so it is not applicable
-for the latter case.
+$8$-connectivity that does not hold for $4$-connectivity, so it is not
+applicable in the latter case.
 
@@
 \texten{
-We will use the fact that for the case of 8-connectivity all $4$ pixels of any
-$2 \times 2$ square are adjacent to each other, and so are contained in the same
-CC and should have same labels upon completion of the labeling algorithm. So
-instead of labeling single pixels we can label $2 \times 2$ groups of pixels, or
-``cells''.
+We will use the fact that in case of 8-connectivity all $4$ pixels of any $2
+\times 2$ square are adjacent to each other, and so are contained in same CC and
+should have same labels upon completion of the labeling algorithm. So instead of
+labeling single pixels we can label $2 \times 2$ groups of pixels, or ``cells''.
 }\textru{
@@
 }
-specified by a binary predicate $b(x, y)$, returning $1$ in case the pixel at
-the given coordinates is black (object pixel), and $0$ otherwise (background
-pixel). For convenience we use 0-indexation throughout the paper, so the topmost
-row has row index $y=0$, and the leftmost column has column index $x=0$.
+specified by binary predicate $b(x, y)$, returning $1$ in case the pixel at the
+given coordinates is an object pixel, and $0$ otherwise (background pixel). For
+convenience we use 0-indexation throughout the paper, so the topmost row is
+associated with row index $y=0$, and the leftmost column is associated with
+column index $x=0$.
 }\textru{
@@
 \texten{
-We need also to define connectivity function for the condensed image.
+We need also to define a connectivity function for the condensed image.
 We cannot actually just use 8-connectivity, as whether two cells are adjacent
@@
 \texten{
-However, for every variant of relative placement of two nearby cells there is a
-mask of pixels of each cells that are important for direct connectivity of this
-pair of cells. So two cells are considered directly connected, if they are
+However, for every possible case of relative placement of two nearby cells there
+is a mask of pixels of each cells that are important for direct connectivity of
+this pair of cells. So two cells are considered directly connected, if they are
 nearby in terms of 8-connectivity and each of them contains at least one object
@@
   \end{subfigure}
-  \begin{subfigure}[t]{0.25\linewidth}
+  \begin{subfigure}[t]{0.2\linewidth}
     \centering
@@
   \label{fig:connectivity:masks}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.475\linewidth}
+    \centering
+    \begin{lstlisting}
+    if (b(x, y)) {
+      if (b(x, y-1)) {
+        // copy labels, or union label sets
+      }
+      // consider other directions
+    }
+    \end{lstlisting}
+    \caption{Binary image with 8-connectivity.}
+  \end{subfigure}
+  \quad
+  \begin{subfigure}[t]{0.475\linewidth}
+    \centering
+    \begin{lstlisting}
+    if (c(x, y)) {
+      if ((c(x, y-1) & ((1<<2) | (1<<3)))
+       && (c(x, y)   & ((1<<0) | (1<<1)))) {
+        // copy labels, or union label sets
+      }
+      // consider other directions
+    }
+    \end{lstlisting}
+    \caption{Condensed image with new connectivity.}
+  \end{subfigure}
+  \caption{Checking the nearby top pixel/cell.}
+  \label{fig:connectivity:code}
 \end{figure}
@@
 that as bitwise operations are usually relatively cheap, we added only a little
-complication compared to decreasing the number of pixels to process by the
+complication compared to the decrease of the number of pixels to process by the
 factor of 4.
@@
 
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{0.475\linewidth}
-    \centering
-    \begin{lstlisting}
-    if (b(x, y)) {
-      if (b(x, y-1)) {
-        // copy labels, or union label sets
-      }
-      // consider other directions
-    }
-    \end{lstlisting}
-    \caption{Binary image with 8-connectivity.}
-  \end{subfigure}
-  \quad
-  \begin{subfigure}[t]{0.475\linewidth}
-    \centering
-    \begin{lstlisting}
-    if (c(x, y)) {
-      if ((c(x, y-1) & ((1<<2) | (1<<3)))
-       && (c(x, y)   & ((1<<0) | (1<<1)))) {
-        // copy labels, or union label sets
-      }
-      // consider other directions
-    }
-    \end{lstlisting}
-    \caption{Condensed image with new connectivity.}
-  \end{subfigure}
-  \caption{Checking the nearby top pixel/cell.}
-  \label{fig:connectivity:code}
-\end{figure}
-
 \section{Benchmarks}
@@
 Also to avoid repeating ``pixel or cell'' we will just use the word ``pixel''
-meaning the appropriate raster for non-condensed and condensed images. When we
-mention neighbors of some pixel, we mean pixels that are directly connected to
-it in terms of appropriate connectivity.
+meaning the appropriate raster for non-condensed and condensed images. When
+mentioning neighbors of some pixel, we mean pixels that are directly connected
+to it in terms of appropriate connectivity.
 }\textru{
@@
 traversal. Graph traversal algorithms are the general approach for CC labeling
-in graphs, and
-DFS is arguably the most simple.
-Although not intended for image labeling and
-suboptimal in most cases, it still has linear in the number of pixels worst-case
-complexity in both time and memory.
+in graphs, and DFS is arguably the most simple.
+Although not intended for image labeling and suboptimal in most cases, it still
+has linear in the number of pixels worst-case complexity in both time and
+memory.
 
@@
 that each object can be easily associated with unique starting set given only
-its coordinates. So memory consumption of SUF is linear in the number of pixels.
+its coordinates. So memory consumption of SUF is linear in the number of pixels
+($O(HW)$), while its amortized time complexity is ``almost'' linear
+($O(HW\alpha(HW))$).
 }\textru{
@@
 consumption is only $O(W)$, which is significant for embedded systems and
-other resource-tight applications.
+other resource-tight applications. The time complexity is same as that of SUF,
+i.\,e. $O(HW\alpha(HW))$.
 }\textru{
@@
 applicable for condensed image, as it exploits similar observations on
-8-connectivity.
+8-connectivity. So this approach is the direct competitor to our optimizations
+of SUF.
 }\textru{
@@
 the number of image lookups.
-% So, according to our benchmark, this combination is efficient only for dense
-% ($>30\%$) noise.
 }\textru{
@@
 
-So benchmarks were run for several values of $d$ to achieve a picture of how the
-speedup depends on the number of object pixels in the image.
+So benchmarks were run for all $d \in \set{2\%, 5\%, 10\%, 15\%, \ldots, 90\%,
+95\%, 98\%}$ to achieve a picture of how the speedup depends on the number of
+object pixels in the image.
 
@@
 overhead of constructing the extra condensed image was included into the total
-run time, since we measured the total time between calling the CC counting
-subroutine, supplied with only the original image, and returning back from
-it.
+run time, since we measured the total run time of the ``black box'' CC counting
+subroutine, supplied with only the original image, so that all the needed
+preparations are done inside the subroutine.
 
@@
 
-The overall benchmark line plot of all algorithms and optimizations is
-presented on \Cref{fig:overall}.
+The overall benchmark plot of average run time for all algorithms and
+optimizations is presented on \Cref{fig:overall}. The actual measured points are
+known for sure only for the mentioned values of d, the points are connected to
+help distinguish groups of points corresponding to same algorithms.
 
@@
 Here we see that while ``2x2 view'' optimization is more memory-efficient than
-``2x2 copy'', it can slow down the method in some situations, and generally it
-does not give as much performance advantage.
+``2x2 copy'', it can slow down the original method in some situations, and
+generally it does not give as much performance advantage.
 
@@
   \begin{tabular}{l|c|c|c|c}
-    Algorithm & worst speedup & worst speedup density & best speedup & best
-    speedup density \\
+              &               & worst speedup &              & best speedup \\
+    Algorithm & worst speedup & density       & best speedup & density \\
     \hline
@@
   \end{tabular}
-  \caption{The worst and the worst speedups of the considered optimizations}
+  \caption{The worst and the best speedups relative to original methods. The
+    shown speedup of DTSUF is relative to SUF.}
   \label{tab:speedup}
@@
 
-The problem of linear variation computing can be reduced to the problem of
-counting connected components in a binary image with eight-connected
-connectivity. This paper presents an efficient way to convert the initial binary
-image with 8-connectivity into a condensed (non-binary) image with new
-connectivity that is 2 times smaller in each dimension than the initial image.
-Most approaches for connected components labeling are still valid for the new
-image. As time and memory consumptions of conventional approaches usually depend
-linearly on the number of pixels, running them on the condensed image can be up
-to 4 times more efficient than running on the initial image. The method is
-essentially a modification of some known raster algorithms for labeling that
-groups pixels into $2 \times 2$ cells. A performance benchmark of the proposed
-method applied to some approaches on noise images is provided. It is shown at
-least up to 2 times more quick results could be received.
+We present an efficient way to convert the initial binary image with
+8-connectivity into a condensed (non-binary) image with new connectivity that is
+2 times smaller in each dimension than the initial image. Most approaches for
+connected components labeling are still valid for the new image. As time and
+memory consumptions of conventional approaches usually depend linearly on the
+number of pixels, running them on the condensed image can be more efficient than
+running on the initial image. The provided benchmark on random images showed
+worst-case speedup factor of $1.7$ for one of our optimizations of the common
+SUF approach to CC labeling. The same algorithm is shown to outperform the
+previously known decision tree optimization of SUF by at least $1.5$ factor for
+image densities less than 80\%.
 

### 2016-04-20 paper/paper.tex
Fix the obvious typo

@@
 densities, slowing down DFS and SUF2 if run on images with densities less than
-38\% and 28\% respectively. Utilizing the latter optimization also does not
+35\% and 25\% respectively. Utilizing the latter optimization also does not
 speed up SUF on images on density approaching 20\%, although it does not slow

### 2016-07-31 paper/paper.tex

@@
 We present an efficient way to convert the initial binary image with
-8-connectivity into a condensed (non-binary) image with new connectivity that is
-2 times smaller in each dimension than the initial image. Most approaches for
-connected components labeling are still valid for the new image. As time and
-memory consumptions of conventional approaches usually depend linearly on the
-number of pixels, running them on the condensed image can be more efficient than
+8-connectivity into a smaller condensed image with new connectivity.
+Most approaches for connected components labeling are still valid
+for the new image. As time and memory complexities of conventional
+approaches usually depend linearly on the number of pixels, running
+them on the condensed image can be more efficient than
 running on the initial image. The provided benchmark on random images showed
@@
 SUF approach to CC labeling. The same algorithm is shown to outperform the
-previously known decision tree optimization of SUF by at least $1.5$ factor for
-image densities less than 80\%.
+previously known decision tree optimization of SUF by at least $1.5$
+factor for images with less than 80\% density.
 
